### 2022-05-30 paper/paper.tex
Add paper template (not sure how bazel works for this case as we might want the pdf file be tracked in git as well).

\documentclass{article}

\begin{document}
Hello! Today!
\end{document}

### 2022-05-30 setup.sh

@@ -25,7 +25,7 @@ function install () {
 
 BINARY_DIR="/usr/local/bin"
 $SUDO apt-get update || exit 2
-$SUDO apt-get install -y clang-format clang-tidy curl git gcc g++ python3 python3-pip npm || exit 3
+$SUDO apt-get install -y clang-format clang-tidy curl git gcc g++ python3 python3-pip npm texlive || exit 3
 install ${BINARY_DIR} bazel "https://github.com/bazelbuild/bazelisk/releases/download/v1.11.0/bazelisk-linux-amd64"
 install ${BINARY_DIR} buildifier "https://github.com/bazelbuild/buildtools/releases/download/5.1.0/buildifier-linux-amd64"
 $SUDO npm install -g prettier

### 2022-05-30 src/client/client.cc
Add framework for testing system end to end.
It doesn't yet work with multi-cohort transactions or with multiple requests in flight at the same time.

@@ -28,7 +28,7 @@ bool Client::TryGetTransactionResults(
     LOG(INFO) << "Still awaiting some results " << response.DebugString();
     return false;
   }
-  LOG(INFO) << "Got final response " << response.DebugString();
+  LOG(INFO) << "Got final response";
   return true;
 }
 

### 2022-05-30 src/client/client_main.cc

@@ -1,12 +1,34 @@
+#include <fcntl.h>
+
+#include <fstream>
 #include <iostream>
 
+#include "absl/container/flat_hash_map.h"
+#include "absl/container/flat_hash_set.h"
+#include "absl/status/statusor.h"
+#include "absl/strings/escaping.h"
+#include "absl/strings/str_cat.h"
 #include "glog/logging.h"
+#include "google/protobuf/io/zero_copy_stream.h"
+#include "google/protobuf/text_format.h"
 #include "grpc/grpc.h"
 #include "grpcpp/create_channel.h"
 #include "grpcpp/security/credentials.h"
 #include "src/client/client.h"
+#include "src/client/client.pb.h"
+#include "src/cohort/start_cohort_server.h"
+#include "src/coordinator/start_coordinator_server.h"
 #include "src/proto/common.pb.h"
 #include "src/proto/coordinator.grpc.pb.h"
+#include "tools/cpp/runfiles/runfiles.h"
+
+constexpr char* kCoordinatorPort = "58000";
+
+struct RunInfo {
+  absl::flat_hash_set<std::string> cohort_addresses;
+  std::vector<std::pair<coordinator::CommitAtomicTransactionRequest, bool>>
+      requests;
+};
 
 coordinator::CommitAtomicTransactionRequest CreateRequest(
     const std::string& client_transaction_id) {
@@ -40,34 +62,127 @@ void AddPutOp(coordinator::CommitAtomicTransactionRequest& request,
       ->set_int64_value(value);
 }
 
-client::ResponseOrStatus PutSingleNamespace(client::Client& client) {
-  coordinator::CommitAtomicTransactionRequest request_put =
-      CreateRequest("put_single");
-  AddPutOp(request_put, "0.0.0.0:50051", "a", 3);
-  return client.CommitSync(request_put);
+void AddRelativePutOp(coordinator::CommitAtomicTransactionRequest& request,
+                      const std::string& address, const std::string& key,
+                      int64_t relative_value) {
+  common::Operation* put_op = request.mutable_transaction()->add_ops();
+  put_op->mutable_namespace_()->set_address(address);
+  put_op->mutable_put()->set_key(key);
+  put_op->mutable_put()
+      ->mutable_value()
+      ->mutable_relative_value()
+      ->mutable_relative_value()
+      ->set_int64_value(relative_value);
 }
 
-client::ResponseOrStatus PutTwoNamespaces(client::Client& client) {
[…]

### 2022-05-30 src/cohort/cohort_server.cc

@@ -51,20 +51,54 @@ absl::Mutex& CohortServer::GetLock(const std::string& key) {
   return *lock;
 }
 
+absl::Status CohortServer::AcquireWholeDbLock(
+    const common::Transaction& transaction, absl::Time presumed_abort_time,
+    internal::TransactionMetadata& txn_metadata) {
+  bool has_put = false;
+  bool has_get = false;
+  for (const common::Operation& op : transaction.ops()) {
+    if (op.has_put()) {
+      has_put = true;
+    }
+    if (op.has_get()) {
+      has_get = true;
+    }
+    if (has_put && has_get) {
+      break;
+    }
+  }
+  if (has_put) {
+    if (!whole_db_mutex_.WriterLockWhenWithDeadline(absl::Condition::kTrue,
+                                                    presumed_abort_time)) {
+      return absl::DeadlineExceededError(
+          "Could not acquire write lock for the DB before the abort deadline");
+    }
+    txn_metadata.has_whole_db_write_lock = true;
+    return absl::OkStatus();
+  }
+  if (has_get) {
+    if (!whole_db_mutex_.ReaderLockWhenWithDeadline(absl::Condition::kTrue,
+                                                    presumed_abort_time)) {
+      return absl::DeadlineExceededError(
+          "Could not acquire read lock for the DB before the abort deadline");
+    }
+    txn_metadata.has_whole_db_read_lock = true;
+  }
+  return absl::OkStatus();
+}
+
 absl::Status CohortServer::AcquireDbLocks(
     const common::Transaction& transaction, absl::Time presumed_abort_time,
     internal::TransactionMetadata& txn_metadata) {
+  if (!txn_metadata.db->SupportsConcurrentWrites()) {
+    return AcquireWholeDbLock(transaction, presumed_abort_time, txn_metadata);
+  }
   absl::flat_hash_set<std::string> write_and_readwrite_keys;
   // Excludes any keys that are written to as well. Otherwise deadlocks could
   // occur if we wait for both the read and write locks.
   absl::flat_hash_set<std::string> readonly_keys;
   for (const common::Operation& op : transaction.ops()) {
     if (op.has_put()) {
-      if (!txn_metadata.db->SupportsConcurrentWrites() &&
-          !txn_metadata.has_whole_db_lock) {
-        whole_db_mutex_.Lock();
[…]

### 2022-05-30 src/cohort/cohort_server_main.cc

@@ -1,17 +1,9 @@
-#include <iostream>
-#include <memory>
 #include <string>
 #include <thread>
 
 #include "absl/flags/flag.h"
 #include "absl/flags/parse.h"
-#include "absl/strings/str_cat.h"
-#include "grpc/grpc.h"
-#include "grpcpp/security/server_credentials.h"
-#include "grpcpp/server.h"
-#include "grpcpp/server_builder.h"
-#include "src/cohort/cohort_server.h"
-#include "src/db/lmdb_database_transaction_adapter.h"
+#include "src/cohort/start_cohort_server.h"
 
 ABSL_FLAG(std::string, port, "50051", "Port to listen to connections on");
 ABSL_FLAG(double, db_thread_ratio, 0.75,
@@ -22,31 +14,13 @@ ABSL_FLAG(std::string, db_txn_response_dir, "/tmp/txn_responses",
           "Directory for persisted transaction responses in case of a crash "
           "while waiting for the blockchain decision");
 
-void RunServer(const std::string& port, uint num_db_threads,
-               const std::string& db_data_dir,
-               const std::string& db_txn_response_dir) {
-  std::string server_address = absl::StrCat("0.0.0.0:", port);
-  cohort::CohortServer service(
-      num_db_threads, db_txn_response_dir, [&db_data_dir]() {
-        return std::make_unique<db::LMDBDatabaseTransactionAdapter>(
-            db_data_dir);
-      });
-
-  grpc::ServerBuilder builder;
-  builder.AddListeningPort(server_address, grpc::InsecureServerCredentials());
-  builder.RegisterService(&service);
-  std::unique_ptr<grpc::Server> server(builder.BuildAndStart());
-  std::cout << "Server listening on " << server_address << std::endl;
-  server->Wait();
-}
-
 int main(int argc, char** argv) {
   absl::ParseCommandLine(argc, argv);
-  RunServer(absl::GetFlag(FLAGS_port),
-            uint(absl::GetFlag(FLAGS_db_thread_ratio) *
-                 std::thread::hardware_concurrency()),
-            absl::GetFlag(FLAGS_db_data_dir),
-            absl::GetFlag(FLAGS_db_txn_response_dir));
+  cohort::RunServer(absl::GetFlag(FLAGS_port),
+                    uint(absl::GetFlag(FLAGS_db_thread_ratio) *
+                         std::thread::hardware_concurrency()),
[…]

### 2022-05-30 src/cohort/start_cohort_server.h

@@ -1,0 +1,14 @@
+#ifndef SRC_COHORT_START_COHORT_SERVER_H_
+
+#define SRC_COHORT_START_COHORT_SERVER_H_
+
+#include <string>
+
+namespace cohort {
+void RunServer(const std::string& port, uint num_db_threads,
+               const std::string& db_data_dir,
+               const std::string& db_txn_response_dir);
+
+}  // namespace cohort
+
+#endif  // SRC_COHORT_START_COHORT_SERVER_H_

### 2022-05-30 src/coordinator/coordinator_server.cc

@@ -188,19 +188,9 @@ void CoordinatorServer::SendCohortPrepareRequests(
       metadata.possibly_sent_to_all_cohorts = true;
     }
 
-    rpcs.push_back(PrepareCohortTransaction(
-        transaction_id, sub_transaction.namespace_, prepare_request, context));
+    PrepareCohortTransaction(transaction_id, sub_transaction.namespace_,
+                             prepare_request, context);
     ++cohort_index;
-  }
-  for (auto &rpc : rpcs) {
-    // TODO(benjmarks22): Retry failed RPCs.
-    // For now the transactions with failed RPCs will abort at the blockchain
-    // at the presumed abort time.
-    cohort::PrepareTransactionResponse response;
-    grpc::Status status = FinishPrepareCohortTransaction(rpc, response);
-    // Without this line, it does not actually send the request to the cohort.
-    LOG(INFO) << response.DebugString() << " " << status.error_code() << " "
-              << status.error_message();
   }
 }
 
@@ -221,18 +211,17 @@ absl::StatusOr<Blockchain::VotingDecision> CoordinatorServer::GetVotingDecision(
 
 absl::Time CoordinatorServer::Now() { return absl::Now(); }
 
-std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-    cohort::PrepareTransactionResponse>>
-CoordinatorServer::PrepareCohortTransaction(
+void CoordinatorServer::PrepareCohortTransaction(
     const std::string &transaction_id, const Namespace &namespace_,
     const cohort::PrepareTransactionRequest &request,
     const grpc::ServerContext &context) {
   metadata_by_transaction_[transaction_id].contexts.push_back(
       ClientContext::FromServerContext(context));
-  return GetCohortStub(namespace_)
-      .AsyncPrepareTransaction(
+  cohort::PrepareTransactionResponse response;
+  GetCohortStub(namespace_)
+      .PrepareTransaction(
           metadata_by_transaction_[transaction_id].contexts.back().get(),
-          request, &completion_queue_);
+          request, &response);
 }
 
 grpc::Status CoordinatorServer::FinishPrepareCohortTransaction(
@@ -300,7 +289,7 @@ grpc::Status CoordinatorServer::UpdateResponseForSingleCohortTransaction(
   if (!status.ok()) {
     return grpc::Status(
         status.error_code(),
-        absl::StrCat("Failed to get cohort results", status.error_message()));
[…]

### 2022-05-30 src/coordinator/coordinator_server.h

@@ -58,12 +58,10 @@ class CoordinatorServer : public Coordinator::Service {
 
  private:
   // The virtual methods are so that they can be mocked out for testing.
-  virtual std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-      cohort::PrepareTransactionResponse>>
-  PrepareCohortTransaction(const std::string &transaction_id,
-                           const common::Namespace &namespace_,
-                           const cohort::PrepareTransactionRequest &request,
-                           const grpc::ServerContext &context);
+  virtual void PrepareCohortTransaction(
+      const std::string &transaction_id, const common::Namespace &namespace_,
+      const cohort::PrepareTransactionRequest &request,
+      const grpc::ServerContext &context);
   virtual grpc::Status FinishPrepareCohortTransaction(
       std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
           cohort::PrepareTransactionResponse>> &async_response,

### 2022-05-30 src/coordinator/coordinator_server_main.cc

@@ -4,32 +4,14 @@
 
 #include "absl/flags/flag.h"
 #include "absl/flags/parse.h"
-#include "absl/strings/str_cat.h"
-#include "grpc/grpc.h"
-#include "grpcpp/security/server_credentials.h"
-#include "grpcpp/server.h"
-#include "grpcpp/server_builder.h"
-#include "src/coordinator/coordinator_server.h"
+#include "absl/time/time.h"
+#include "src/coordinator/start_coordinator_server.h"
 
 ABSL_FLAG(std::string, port, "50052", "Port to listen to connections on");
 ABSL_FLAG(
-    std::string, default_presumed_abort_duration, "5s",
+    std::string, default_presumed_abort_duration, "1m",
     "Default duration for the presumed abort time relative to the current "
     "time. Only used if the client does not specify the timestamp.");
-
-void RunServer(const std::string& port,
-               absl::Duration default_presumed_abort_duration) {
-  std::string server_address = absl::StrCat("0.0.0.0:", port);
-
-  coordinator::CoordinatorServer service(default_presumed_abort_duration);
-
-  grpc::ServerBuilder builder;
-  builder.AddListeningPort(server_address, grpc::InsecureServerCredentials());
-  builder.RegisterService(&service);
-  std::unique_ptr<grpc::Server> server(builder.BuildAndStart());
-  std::cout << "Server listening on " << server_address << std::endl;
-  server->Wait();
-}
 
 int main(int argc, char** argv) {
   absl::ParseCommandLine(argc, argv);
@@ -37,7 +19,7 @@ int main(int argc, char** argv) {
   std::string error;
   if (absl::ParseFlag(absl::GetFlag(FLAGS_default_presumed_abort_duration),
                       &duration, &error)) {
-    RunServer(absl::GetFlag(FLAGS_port), duration);
+    coordinator::RunServer(absl::GetFlag(FLAGS_port), duration);
   } else {
     std::fprintf(
         stderr,

### 2022-05-30 src/coordinator/coordinator_server_test.cc

@@ -26,14 +26,10 @@ class CoordinatorWithMockCohorts : public coordinator::CoordinatorServer {
       : coordinator::CoordinatorServer(default_presumed_abort_duration) {}
 
   MOCK_METHOD4(MockPrepareCohortTransaction,
-               std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-                   cohort::PrepareTransactionResponse>>(
-                   const std::string& transaction_id,
-                   const common::Namespace& namespace_,
-                   const cohort::PrepareTransactionRequest& request,
-                   const grpc::ServerContext& context));
-  MOCK_METHOD1(MockFinishPrepareCohortTransaction,
-               grpc::Status(cohort::PrepareTransactionResponse& response));
+               void(const std::string& transaction_id,
+                    const common::Namespace& namespace_,
+                    const cohort::PrepareTransactionRequest& request,
+                    const grpc::ServerContext& context));
   MOCK_METHOD1(MockFinishAsyncGetResultsFromCohort,
                grpc::Status(cohort::GetTransactionResultResponse& response));
   MOCK_METHOD4(MockGetResultsFromCohort,
@@ -54,20 +50,11 @@ class CoordinatorWithMockCohorts : public coordinator::CoordinatorServer {
  private:
   bool SortCohortRequests() override { return true; }
   absl::Time Now() override { return MockNow(); }
-  std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-      cohort::PrepareTransactionResponse>>
-  PrepareCohortTransaction(const std::string& transaction_id,
-                           const common::Namespace& namespace_,
-                           const cohort::PrepareTransactionRequest& request,
-                           const grpc::ServerContext& context) override {
-    return MockPrepareCohortTransaction(transaction_id, namespace_, request,
-                                        context);
-  }
-  grpc::Status FinishPrepareCohortTransaction(
-      std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-          cohort::PrepareTransactionResponse>>& /*async_response*/,
-      cohort::PrepareTransactionResponse& response) override {
-    return MockFinishPrepareCohortTransaction(response);
+  void PrepareCohortTransaction(
+      const std::string& transaction_id, const common::Namespace& namespace_,
+      const cohort::PrepareTransactionRequest& request,
+      const grpc::ServerContext& context) override {
[…]

### 2022-05-30 src/coordinator/start_coordinator_server.cc

@@ -1,0 +1,28 @@
+#include "src/coordinator/start_coordinator_server.h"
+
+#include <iostream>
+#include <memory>
+
+#include "absl/strings/str_cat.h"
+#include "grpc/grpc.h"
+#include "grpcpp/security/server_credentials.h"
+#include "grpcpp/server.h"
+#include "grpcpp/server_builder.h"
+#include "src/coordinator/coordinator_server.h"
+
+namespace coordinator {
+
+void RunServer(const std::string& port,
+               absl::Duration default_presumed_abort_duration) {
+  std::string server_address = absl::StrCat("0.0.0.0:", port);
+
+  CoordinatorServer service(default_presumed_abort_duration);
+
+  grpc::ServerBuilder builder;
+  builder.AddListeningPort(server_address, grpc::InsecureServerCredentials());
+  builder.RegisterService(&service);
+  std::unique_ptr<grpc::Server> server(builder.BuildAndStart());
+  std::cout << "Server listening on " << server_address << std::endl;
+  server->Wait();
+}
+}  // namespace coordinator

### 2022-05-31 src/blockchain/two_phase_commit.cc
Integrate blockchain adapter with C++.
I'm still working on making the adapter work with Bazel, but in the meantime it should be possible to test end to end with a combination of Bazel and node.

@@ -1,25 +1,49 @@
 #include "src/blockchain/two_phase_commit.h"
 
+#include "src/utils/status_utils.h"
+
 namespace blockchain {
 
-absl::Status TwoPhaseCommit::StartVoting(const std::string& /*transaction_id*/,
-                                         const std::time_t& /*timeout_time*/,
-                                         int /*n_participants*/) {
-  // TODO(heronyang): Implement this.
-  return absl::UnimplementedError("Unimplemented.");
+absl::Status TwoPhaseCommit::StartVoting(const std::string& transaction_id,
+                                         const std::time_t& timeout_time,
+                                         int n_participants) {
+  grpc::ClientContext context;
+  StartVotingRequest request;
+  request.set_transaction_id(transaction_id);
+  const absl::Time absl_time = absl::FromTimeT(timeout_time);
+  timespec ts = absl::ToTimespec(absl_time);
+  google::protobuf::Timestamp timestamp;
+  request.mutable_timeout_time()->set_seconds(ts.tv_sec);
+  request.mutable_timeout_time()->set_nanos(ts.tv_nsec);
+  request.set_cohorts(n_participants);
+  StartVotingResponse response;
+  grpc::Status status = stub_->StartVoting(&context, request, &response);
+  return utils::FromGrpcStatus(status, "Failed to start voting");
 }
 
-absl::Status TwoPhaseCommit::Vote(const std::string& /*transaction_id*/,
-                                  int /*participant_id*/,
-                                  TwoPhaseCommit::Ballot /*ballot*/) {
-  // TODO(heronyang): Implement this.
-  return absl::UnimplementedError("Unimplemented.");
+absl::Status TwoPhaseCommit::Vote(const std::string& transaction_id,
+                                  int participant_id, Ballot ballot) {
+  grpc::ClientContext context;
+  VoteRequest request;
+  request.set_transaction_id(transaction_id);
+  request.set_cohort_id(participant_id);
+  request.set_ballot(ballot);
+  VoteResponse response;
+  grpc::Status status = stub_->Vote(&context, request, &response);
+  return utils::FromGrpcStatus(status, "Failed to vote");
 }
 
-absl::StatusOr<TwoPhaseCommit::VotingDecision>
-TwoPhaseCommit::GetVotingDecision(const std::string& /*transaction_id*/) {
-  // TODO(heronyang): Implement this.
-  return absl::UnimplementedError("Unimplemented.");
[…]

### 2022-05-31 src/blockchain/two_phase_commit.h

@@ -3,27 +3,38 @@
 #define SRC_BLOCKCHAIN_TWO_PHASE_COMMIT_H_
 
 #include <ctime>
+#include <memory>
 #include <string>
 
 #include "absl/status/status.h"
 #include "absl/status/statusor.h"
+#include "grpcpp/channel.h"
+#include "grpcpp/client_context.h"
+#include "src/blockchain/proto/two_phase_commit_adapter.grpc.pb.h"
 
 namespace blockchain {
 class TwoPhaseCommit {
  public:
-  static absl::Status StartVoting(const std::string &transaction_id,
-                                  const std::time_t &timeout_time,
-                                  int n_participants);
+  explicit TwoPhaseCommit(std::shared_ptr<grpc::Channel> channel)
+      : TwoPhaseCommit(TwoPhaseCommitAdapter::NewStub(channel)) {}
 
-  enum class Ballot { ABORT, COMMIT };
-  static absl::Status Vote(const std::string &transaction_id,
-                           int participant_id, Ballot ballot);
+  explicit TwoPhaseCommit(
+      std::unique_ptr<TwoPhaseCommitAdapter::StubInterface> stub)
+      : stub_(stub.release()) {}
 
-  enum class VotingDecision { PENDING, ABORT, COMMIT };
+  absl::Status StartVoting(const std::string &transaction_id,
+                           const std::time_t &timeout_time, int n_participants);
+
+  absl::Status Vote(const std::string &transaction_id, int participant_id,
+                    Ballot ballot);
+
   // Gets the voting decision of a transaction.
   // TODO(heronyang): Attach ABORT_REASON, PENDING_REASON to the response.
-  static absl::StatusOr<VotingDecision> GetVotingDecision(
+  absl::StatusOr<VotingDecision> GetVotingDecision(
       const std::string &transaction_id);
+
+ private:
+  std::unique_ptr<TwoPhaseCommitAdapter::StubInterface> stub_;
 };
 
 }  // namespace blockchain

### 2022-05-31 src/cohort/cohort_server.cc

@@ -18,22 +18,6 @@ namespace {
 using Blockchain = ::blockchain::TwoPhaseCommit;
 using ::grpc::ServerContext;
 
-Blockchain::VotingDecision WaitForBlockchainDecision(
-    const std::string& transaction_id, absl::Time presumed_abort_time) {
-  std::this_thread::sleep_until(absl::ToChronoTime(presumed_abort_time));
-  Blockchain::VotingDecision decision =
-      Blockchain::GetVotingDecision(transaction_id)
-          .value_or(Blockchain::VotingDecision::PENDING);
-  while (decision == Blockchain::VotingDecision::PENDING) {
-    VLOG(1) << "Waiting for blockchain decision";
-    std::this_thread::sleep_for(
-        absl::ToChronoMicroseconds(absl::Now() - presumed_abort_time));
-    decision = Blockchain::GetVotingDecision(transaction_id)
-                   .value_or(Blockchain::VotingDecision::PENDING);
-  }
-  return decision;
-}
-
 template <typename Message>
 bool WriteToFile(const Message& message, const std::string& path) {
   std::fstream output(path, std::ios::out | std::ios::trunc | std::ios::binary);
@@ -143,6 +127,28 @@ absl::Status CohortServer::ProcessTransactionInDb(
   return absl::OkStatus();
 }
 
+blockchain::VotingDecision CohortServer::WaitForBlockchainDecision(
+    const std::string& transaction_id, absl::Time presumed_abort_time) {
+  bool waited_until_presumed_abort_time = false;
+  while (true) {
+    const blockchain::VotingDecision decision =
+        blockchain_->GetVotingDecision(transaction_id)
+            .value_or(blockchain::VotingDecision::VOTING_DECISION_PENDING);
+    if (decision != blockchain::VotingDecision::VOTING_DECISION_PENDING &&
+        decision != blockchain::VotingDecision::VOTING_DECISION_UNKNOWN) {
+      return decision;
+    }
+    VLOG(1) << "Waiting for blockchain decision";
+    if (waited_until_presumed_abort_time) {
+      std::this_thread::sleep_for(
+          absl::ToChronoMicroseconds(absl::Now() - presumed_abort_time));
+    } else {
+      std::this_thread::sleep_until(absl::ToChronoTime(presumed_abort_time));
+      waited_until_presumed_abort_time = true;
+    }
+  }
+}
+
 void CohortServer::AbortTransaction(const std::string& transaction_id,
                                     int cohort_index,
[…]

### 2022-05-31 src/cohort/cohort_server.h

@@ -8,6 +8,7 @@
 #include "absl/container/flat_hash_set.h"
 #include "absl/status/statusor.h"
 #include "grpcpp/server_context.h"
+#include "src/blockchain/two_phase_commit.h"
 #include "src/db/database_transaction_adapter.h"
 #include "src/proto/cohort.grpc.pb.h"
 #include "thread_pool.hpp"
@@ -28,10 +29,12 @@ class CohortServer : public Cohort::Service {
  public:
   CohortServer(uint num_db_threads, const std::string& db_txn_response_dir,
                std::function<std::unique_ptr<db::DatabaseTransactionAdapter>()>
-                   db_transaction_adapter_creator)
+                   db_transaction_adapter_creator,
+               std::unique_ptr<blockchain::TwoPhaseCommit> blockchain)
       : thread_pool_(num_db_threads),
         db_txn_response_dir_(db_txn_response_dir),
-        db_transaction_adapter_creator_(db_transaction_adapter_creator) {}
+        db_transaction_adapter_creator_(db_transaction_adapter_creator),
+        blockchain_(blockchain.release()) {}
 
   grpc::Status PrepareTransaction(
       grpc::ServerContext* context, const PrepareTransactionRequest* request,
@@ -49,6 +52,9 @@ class CohortServer : public Cohort::Service {
                               internal::TransactionMetadata& txn_metadata);
 
   void ProcessTransaction(const PrepareTransactionRequest& request);
+
+  blockchain::VotingDecision WaitForBlockchainDecision(
+      const std::string& transaction_id, absl::Time presumed_abort_time);
 
   void AbortTransaction(const std::string& transaction_id, int cohort_index,
                         absl::optional<absl::Status> abort_status);
@@ -81,6 +87,7 @@ class CohortServer : public Cohort::Service {
   absl::flat_hash_map<std::string, std::unique_ptr<absl::Mutex>> locks_by_key_;
   // Used for DBs that don't support concurrent write transactions.
   absl::Mutex whole_db_mutex_;
+  std::unique_ptr<blockchain::TwoPhaseCommit> blockchain_;
 };
 
 }  // namespace cohort

### 2022-05-31 src/cohort/cohort_server_main.cc

@@ -7,13 +7,17 @@
 #include "absl/flags/parse.h"
 #include "absl/strings/str_cat.h"
 #include "grpc/grpc.h"
-#include "grpcpp/security/server_credentials.h"
+#include "grpcpp/create_channel.h"
+#include "grpcpp/security/credentials.h"
 #include "grpcpp/server.h"
 #include "grpcpp/server_builder.h"
+#include "src/blockchain/two_phase_commit.h"
 #include "src/cohort/cohort_server.h"
 #include "src/db/lmdb_database_transaction_adapter.h"
 
 ABSL_FLAG(std::string, port, "50051", "Port to listen to connections on");
+ABSL_FLAG(std::string, blockchain_adapter_port, "50551",
+          "Port to listen to the blockchain adapter server on");
 ABSL_FLAG(double, db_thread_ratio, 0.75,
           "Fraction of threads to assign to DB processing. The remaining "
           "threads are used by gRPC.");
@@ -22,15 +26,21 @@ ABSL_FLAG(std::string, db_txn_response_dir, "/tmp/txn_responses",
           "Directory for persisted transaction responses in case of a crash "
           "while waiting for the blockchain decision");
 
-void RunServer(const std::string& port, uint num_db_threads,
+void RunServer(const std::string& port,
+               const std::string& blockchain_adapter_port, uint num_db_threads,
                const std::string& db_data_dir,
                const std::string& db_txn_response_dir) {
   std::string server_address = absl::StrCat("0.0.0.0:", port);
+  std::string blockchain_adapter_address =
+      absl::StrCat("0.0.0.0:", blockchain_adapter_port);
   cohort::CohortServer service(
-      num_db_threads, db_txn_response_dir, [&db_data_dir]() {
+      num_db_threads, db_txn_response_dir,
+      [&db_data_dir]() {
         return std::make_unique<db::LMDBDatabaseTransactionAdapter>(
             db_data_dir);
-      });
+      },
+      std::make_unique<blockchain::TwoPhaseCommit>(grpc::CreateChannel(
+          blockchain_adapter_address, grpc::InsecureChannelCredentials())));
 
   grpc::ServerBuilder builder;
   builder.AddListeningPort(server_address, grpc::InsecureServerCredentials());
@@ -43,6 +53,7 @@ void RunServer(const std::string& port, uint num_db_threads,
 int main(int argc, char** argv) {
[…]

### 2022-05-31 src/cohort/start_cohort_with_blockchain_adapter_server.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+set -eEuo pipefail
+
+# --- begin runfiles.bash initialization ---
+if [[ ! -d "${RUNFILES_DIR:-/dev/null}" && ! -f "${RUNFILES_MANIFEST_FILE:-/dev/null}" ]]; then
+    if [[ -f "$0.runfiles_manifest" ]]; then
+      export RUNFILES_MANIFEST_FILE="$0.runfiles_manifest"
+    elif [[ -f "$0.runfiles/MANIFEST" ]]; then
+      export RUNFILES_MANIFEST_FILE="$0.runfiles/MANIFEST"
+    elif [[ -f "$0.runfiles/bazel_tools/tools/bash/runfiles/runfiles.bash" ]]; then
+      export RUNFILES_DIR="$0.runfiles"
+    fi
+fi
+if [[ -f "${RUNFILES_DIR:-/dev/null}/bazel_tools/tools/bash/runfiles/runfiles.bash" ]]; then
+  source "${RUNFILES_DIR}/bazel_tools/tools/bash/runfiles/runfiles.bash"
+elif [[ -f "${RUNFILES_MANIFEST_FILE:-/dev/null}" ]]; then
+  source "$(grep -m1 "^bazel_tools/tools/bash/runfiles/runfiles.bash " \
+            "$RUNFILES_MANIFEST_FILE" | cut -d ' ' -f 2-)"
+else
+  echo >&2 "ERROR: cannot find @bazel_tools//tools/bash/runfiles:runfiles.bash"
+  exit 1
+fi
+# --- end runfiles.bash initialization ---
+
+contract_address=$4
+adapter_server_port=${5:-61362}
+cohort_port=${6:-51000}
+# Don't change this since it's the first one in the deterministic ganache client.
+account=${7:-'0x90F8bf6A479f320ead074411a4B0e7944Ea8c9C1'}
+blockchain_port=${8:-7545}
+db_data_dir=${9:-'/tmp/data'}
+db_txn_response_dir=${10:-'/tmp/txn_responses'}
+db_thread_ratio=${11:-0.75}
+$(rlocation __main__/src/blockchain/client/run_adapter_server.sh) $contract_address $adapter_server_port $account $blockchain_port &
+sleep 3s
[…]

### 2022-05-31 src/coordinator/coordinator_server.cc

@@ -21,7 +21,6 @@ namespace coordinator {
 
 namespace {
 
-using Blockchain = ::blockchain::TwoPhaseCommit;
 using CohortStub = ::cohort::Cohort::StubInterface;
 using ::common::Namespace;
 using ::coordinator::internal::SubTransaction;
@@ -198,12 +197,12 @@ absl::Status CoordinatorServer::StartVoting(
   const time_t abort_time =
       absl::ToTimeT(absl::FromUnixSeconds(presumed_abort_time.seconds()) +
                     absl::Nanoseconds(presumed_abort_time.nanos()));
-  return Blockchain::StartVoting(transaction_id, abort_time, num_cohorts);
+  return blockchain_->StartVoting(transaction_id, abort_time, num_cohorts);
 }
 
-absl::StatusOr<Blockchain::VotingDecision> CoordinatorServer::GetVotingDecision(
+absl::StatusOr<blockchain::VotingDecision> CoordinatorServer::GetVotingDecision(
     const std::string &transaction_id) {
-  return Blockchain::GetVotingDecision(transaction_id);
+  return blockchain_->GetVotingDecision(transaction_id);
 }
 
 absl::Time CoordinatorServer::Now() { return absl::Now(); }
@@ -379,7 +378,10 @@ grpc::Status CoordinatorServer::GetTransactionResult(
     }
     return status;
   }
-  if (metadata.decision == Blockchain::VotingDecision::PENDING) {
+  if (metadata.decision ==
+          blockchain::VotingDecision::VOTING_DECISION_PENDING ||
+      metadata.decision ==
+          blockchain::VotingDecision::VOTING_DECISION_UNKNOWN) {
     const auto decision_or_status =
         GetVotingDecision(request->global_transaction_id());
     if (!decision_or_status.ok()) {
@@ -389,16 +391,17 @@ grpc::Status CoordinatorServer::GetTransactionResult(
     metadata.decision = decision_or_status.value();
   }
   switch (metadata.decision) {
-    case Blockchain::VotingDecision::PENDING:
+    case blockchain::VotingDecision::VOTING_DECISION_UNKNOWN:
+    case blockchain::VotingDecision::VOTING_DECISION_PENDING:
       response->mutable_pending_response();
       break;
-    case Blockchain::VotingDecision::ABORT:
+    case blockchain::VotingDecision::VOTING_DECISION_ABORT:
       UpdateAbortedResponseFromCohorts(request->global_transaction_id(),
                                        *context);
       *response =
[…]

### 2022-05-31 src/coordinator/coordinator_server.h

@@ -24,7 +24,7 @@ struct TransactionMetadata {
   // This indicates that the transaction may have been sent to all the cohorts
   // and thus might commit.
   bool possibly_sent_to_all_cohorts;
-  blockchain::TwoPhaseCommit::VotingDecision decision;
+  blockchain::VotingDecision decision;
   absl::flat_hash_set<std::string> cohorts_already_responded;
   // TODO(benjmarks22): Add locks so that only one thread can access this data
   // at once.
@@ -44,8 +44,11 @@ struct SubTransaction {
 
 class CoordinatorServer : public Coordinator::Service {
  public:
-  explicit CoordinatorServer(absl::Duration default_presumed_abort_duration)
-      : default_presumed_abort_duration_(default_presumed_abort_duration) {}
+  explicit CoordinatorServer(
+      absl::Duration default_presumed_abort_duration,
+      std::unique_ptr<blockchain::TwoPhaseCommit> blockchain)
+      : default_presumed_abort_duration_(default_presumed_abort_duration),
+        blockchain_(blockchain.release()) {}
 
   grpc::Status CommitAtomicTransaction(
       grpc::ServerContext *context,
@@ -81,8 +84,8 @@ class CoordinatorServer : public Coordinator::Service {
       const std::string &transaction_id,
       const google::protobuf::Timestamp &presumed_abort_time,
       size_t num_cohorts);
-  virtual absl::StatusOr<blockchain::TwoPhaseCommit::VotingDecision>
-  GetVotingDecision(const std::string &transaction_id);
+  virtual absl::StatusOr<blockchain::VotingDecision> GetVotingDecision(
+      const std::string &transaction_id);
   virtual cohort::Cohort::StubInterface &GetCohortStub(
       const common::Namespace &namespace_);
   virtual bool SortCohortRequests() { return false; }
@@ -113,6 +116,7 @@ class CoordinatorServer : public Coordinator::Service {
   absl::flat_hash_map<std::string, internal::TransactionResponse>
       response_by_transaction_;
   absl::Duration default_presumed_abort_duration_;
+  std::unique_ptr<blockchain::TwoPhaseCommit> blockchain_;
 };
 
 }  // namespace coordinator

### 2022-05-31 src/coordinator/coordinator_server_main.cc

@@ -6,22 +6,33 @@
 #include "absl/flags/parse.h"
 #include "absl/strings/str_cat.h"
 #include "grpc/grpc.h"
+#include "grpcpp/create_channel.h"
+#include "grpcpp/security/credentials.h"
 #include "grpcpp/security/server_credentials.h"
 #include "grpcpp/server.h"
 #include "grpcpp/server_builder.h"
+#include "src/blockchain/two_phase_commit.h"
 #include "src/coordinator/coordinator_server.h"
 
 ABSL_FLAG(std::string, port, "50052", "Port to listen to connections on");
+ABSL_FLAG(std::string, blockchain_adapter_port, "50552",
+          "Port to listen to the blockchain adapter server on");
 ABSL_FLAG(
     std::string, default_presumed_abort_duration, "1m",
     "Default duration for the presumed abort time relative to the current "
     "time. Only used if the client does not specify the timestamp.");
 
 void RunServer(const std::string& port,
+               const std::string& blockchain_adapter_port,
                absl::Duration default_presumed_abort_duration) {
   std::string server_address = absl::StrCat("0.0.0.0:", port);
+  std::string blockchain_adapter_address =
+      absl::StrCat("0.0.0.0:", blockchain_adapter_port);
 
-  coordinator::CoordinatorServer service(default_presumed_abort_duration);
+  coordinator::CoordinatorServer service(
+      default_presumed_abort_duration,
+      std::make_unique<blockchain::TwoPhaseCommit>(grpc::CreateChannel(
+          blockchain_adapter_address, grpc::InsecureChannelCredentials())));
 
   grpc::ServerBuilder builder;
   builder.AddListeningPort(server_address, grpc::InsecureServerCredentials());
@@ -37,7 +48,8 @@ int main(int argc, char** argv) {
   std::string error;
   if (absl::ParseFlag(absl::GetFlag(FLAGS_default_presumed_abort_duration),
                       &duration, &error)) {
-    RunServer(absl::GetFlag(FLAGS_port), duration);
+    RunServer(absl::GetFlag(FLAGS_port),
+              absl::GetFlag(FLAGS_blockchain_adapter_port), duration);
   } else {
     std::fprintf(
         stderr,

### 2022-05-31 src/coordinator/coordinator_server_test.cc

@@ -5,6 +5,7 @@
 #include "grpcpp/server_context.h"
 #include "gtest/gtest.h"
 #include "protobuf-matchers/protocol-buffer-matchers.h"
+#include "src/blockchain/proto/two_phase_commit_adapter_mock.grpc.pb.h"
 #include "src/proto/cohort_mock.grpc.pb.h"
 #include "src/proto/common.pb.h"
 #include "src/proto/coordinator.pb.h"
@@ -23,7 +24,11 @@ class CoordinatorWithMockCohorts : public coordinator::CoordinatorServer {
  public:
   explicit CoordinatorWithMockCohorts(
       absl::Duration default_presumed_abort_duration)
-      : coordinator::CoordinatorServer(default_presumed_abort_duration) {}
+      : coordinator::CoordinatorServer(
+            default_presumed_abort_duration,
+            std::make_unique<blockchain::TwoPhaseCommit>(
+                std::make_unique<
+                    blockchain::MockTwoPhaseCommitAdapterStub>())) {}
 
   MOCK_METHOD4(MockPrepareCohortTransaction,
                void(const std::string& transaction_id,
@@ -43,7 +48,7 @@ class CoordinatorWithMockCohorts : public coordinator::CoordinatorServer {
                    const google::protobuf::Timestamp& presumed_abort_time,
                    size_t num_cohorts));
   MOCK_METHOD1(MockGetVotingDecision,
-               absl::StatusOr<blockchain::TwoPhaseCommit::VotingDecision>(
+               absl::StatusOr<blockchain::VotingDecision>(
                    const std::string& transaction_id));
   MOCK_METHOD0(MockNow, absl::Time());
 
@@ -83,7 +88,7 @@ class CoordinatorWithMockCohorts : public coordinator::CoordinatorServer {
       size_t num_cohorts) override {
     return MockStartVoting(transaction_id, presumed_abort_time, num_cohorts);
   }
-  absl::StatusOr<blockchain::TwoPhaseCommit::VotingDecision> GetVotingDecision(
+  absl::StatusOr<blockchain::VotingDecision> GetVotingDecision(
       const std::string& transaction_id) override {
     return MockGetVotingDecision(transaction_id);
   }
@@ -363,7 +368,7 @@ TEST(CoordinatorServerTest, GetsResultsWhenCommitted) {
       commit_response.global_transaction_id());
[…]

### 2022-05-31 src/utils/status_utils.cc

@@ -10,4 +10,10 @@ grpc::Status FromAbslStatus(const absl::Status& status,
   return grpc::Status(static_cast<grpc::StatusCode>(status.raw_code()),
                       absl::StrCat(prefix, std::string(status.message())));
 }
+absl::Status FromGrpcStatus(const grpc::Status& status,
+                            const std::string& prefix) {
+  return absl::Status(
+      static_cast<absl::StatusCode>(status.error_code()),
+      absl::StrCat(prefix, std::string(status.error_message())));
+}
 }  // namespace utils

### 2022-05-31 src/utils/status_utils.h

@@ -8,6 +8,8 @@
 namespace utils {
 grpc::Status FromAbslStatus(const absl::Status& status,
                             const std::string& prefix);
+absl::Status FromGrpcStatus(const grpc::Status& status,
+                            const std::string& prefix);
 
 #define ASSIGN_OR_RETURN(lhs, rexpr)         \
   auto result_assign_or_return = rexpr;      \

### 2022-05-31 src/blockchain/client/adapter_server.js
Add script to run blockchain adapter with cohort and coordinator

@@ -3,10 +3,12 @@ const args = process.argv.slice(2);
 const blockchainUri = 'ws://localhost:' + args[0];
 const contractAddr = args[1];
 const sharedAccount = args[2];
+const port = args[3];
 
 console.log('Blockchain URI', blockchainUri);
 console.log('Contract Address', contractAddr);
 console.log('Shared Account', sharedAccount);
+console.log('Port', port);
 
 const PROTO_PATH = __dirname + '/../proto/two_phase_commit_adapter.proto';
 
@@ -24,7 +26,8 @@ var twoPhaseCommitAdapterProto =
 
 var TwoPhaseCommitClient = require('./contract_client.js');
 var contractClient = new TwoPhaseCommitClient(
-    '../build/contracts/TwoPhaseCommit.json', blockchainUri, contractAddr);
+    'src/blockchain/build/contracts/TwoPhaseCommit.json', blockchainUri,
+    contractAddr);
 
 function startVoting(call, callback) {
   console.log('Received: startVoting', call.request);
@@ -90,9 +93,11 @@ function main() {
     getVotingDecision: getVotingDecision,
     getHeartBeat: getHeartBeat
   });
+  console.log('Starting server')
   server.bindAsync(
-      '0.0.0.0:50051', grpc.ServerCredentials.createInsecure(), () => {
+      '0.0.0.0:' + port, grpc.ServerCredentials.createInsecure(), () => {
         server.start();
+        console.log('Started server');
       });
 }
 

### 2022-05-31 src/blockchain/client/run_adapter_server.sh

@@ -20,7 +20,30 @@ else
   exit 1
 fi
 # --- end runfiles.bash initialization ---
-
-setup_truffle="$(rlocation __main__/src/blockchain/setup_truffle.sh)"
-source $setup_truffle
-$truffle compile
+compile_script=$(rlocation __main__/src/blockchain/compile_command.bash)
+$compile_script $1 $2
+shift
+shift
+adapter_server=$(rlocation __main__/src/blockchain/client/adapter_server_command.bash)
+contract_address=''
+adapter_server_port=61362
+# Don't change this since it's the first one in the deterministic ganache client.
+account='0x90F8bf6A479f320ead074411a4B0e7944Ea8c9C1'
+blockchain_port=7545
+extra_args=(" ")
+while [ $# -gt 0 ]
+do
+  flag=${1%%=*}
+  value="${1#*=}"
+  echo $flag
+  echo $value
+  case "$flag" in
+    --contract_address*) contract_address="$value";;
+    --adapter_server_port*) adapter_server_port="$value";;
+    --account*) account="$value";;
+    --blockchain_port*) blockchain_port="$value";;
+    *) extra_args+=" $1";;
+  esac
+  shift
+done
+$adapter_server $blockchain_port $contract_address $account $adapter_server_port $extra_args{@} &

### 2022-05-31 src/blockchain/two_phase_commit.cc

@@ -46,4 +46,12 @@ absl::StatusOr<VotingDecision> TwoPhaseCommit::GetVotingDecision(
   return response.decision();
 }
 
+absl::Status TwoPhaseCommit::GetHeartBeat() {
+  grpc::ClientContext context;
+  GetHeartBeatRequest request;
+  GetHeartBeatResponse response;
+  grpc::Status status = stub_->GetHeartBeat(&context, request, &response);
+  return utils::FromGrpcStatus(status, "Failed to get heartbeat");
+}
+
 }  // namespace blockchain

### 2022-05-31 src/blockchain/two_phase_commit.h

@@ -5,9 +5,12 @@
 #include <ctime>
 #include <memory>
 #include <string>
+#include <thread>
 
 #include "absl/status/status.h"
 #include "absl/status/statusor.h"
+#include "absl/time/time.h"
+#include "glog/logging.h"
 #include "grpcpp/channel.h"
 #include "grpcpp/client_context.h"
 #include "src/blockchain/proto/two_phase_commit_adapter.grpc.pb.h"
@@ -16,7 +19,12 @@ namespace blockchain {
 class TwoPhaseCommit {
  public:
   explicit TwoPhaseCommit(std::shared_ptr<grpc::Channel> channel)
-      : TwoPhaseCommit(TwoPhaseCommitAdapter::NewStub(channel)) {}
+      : TwoPhaseCommit(TwoPhaseCommitAdapter::NewStub(channel)) {
+    while (!GetHeartBeat().ok()) {
+      LOG(INFO) << "Failed to get heartbeat. Waiting another second.";
+      std::this_thread::sleep_for(absl::ToChronoSeconds(absl::Seconds(1)));
+    }
+  }
 
   explicit TwoPhaseCommit(
       std::unique_ptr<TwoPhaseCommitAdapter::StubInterface> stub)
@@ -32,6 +40,8 @@ class TwoPhaseCommit {
   // TODO(heronyang): Attach ABORT_REASON, PENDING_REASON to the response.
   absl::StatusOr<VotingDecision> GetVotingDecision(
       const std::string &transaction_id);
+
+  absl::Status GetHeartBeat();
 
  private:
   std::unique_ptr<TwoPhaseCommitAdapter::StubInterface> stub_;

### 2022-05-31 src/cohort/start_cohort_with_blockchain_adapter_server.sh

@@ -22,15 +22,25 @@ else
 fi
 # --- end runfiles.bash initialization ---
 
-contract_address=$4
-adapter_server_port=${5:-61362}
-cohort_port=${6:-51000}
-# Don't change this since it's the first one in the deterministic ganache client.
-account=${7:-'0x90F8bf6A479f320ead074411a4B0e7944Ea8c9C1'}
-blockchain_port=${8:-7545}
-db_data_dir=${9:-'/tmp/data'}
-db_txn_response_dir=${10:-'/tmp/txn_responses'}
-db_thread_ratio=${11:-0.75}
-$(rlocation __main__/src/blockchain/client/run_adapter_server.sh) $contract_address $adapter_server_port $account $blockchain_port &
-sleep 3s
-$(rlocation __main__/src/cohort/cohort_server_main) --port=$cohort_port --blockchain_adapter_port=$adapter_server_port --db_data_dir=$db_data_dir --db_txn_response_dir=$db_txn_response_dir --db_thread_ratio=$db_thread_ratio
+adapter_server_args=()
+cohort_server_args=()
+# Needs to be consistent between JS adapter server and cohort server.
+adapter_server_port=51560
+while [ $# -gt 0 ]
+do
+  case "$1" in
+    --adapter_server_port*) adapter_server_port=${1#*=};;
+    --) shift; break;;
+    *) adapter_server_args+=" $1";;
+  esac
+  shift
+done
+while [ $# -gt 0 ]
+do
+  cohort_server_args+=" $1"
+  shift
+done
+adapter_server_args+=" --adapter_server_port=$adapter_server_port"
+cohort_server_args+=" --blockchain_adapter_port=$adapter_server_port"
+$(rlocation __main__/src/blockchain/client/run_adapter_server_command.bash) $adapter_server_args
+$(rlocation __main__/src/cohort/cohort_server_main) $cohort_server_args

### 2022-05-31 src/coordinator/start_coordinator_with_blockchain_adapter_server.sh

@@ -1,5 +1,6 @@
 #!/bin/bash
 set -eEuo pipefail
+
 # --- begin runfiles.bash initialization ---
 if [[ ! -d "${RUNFILES_DIR:-/dev/null}" && ! -f "${RUNFILES_MANIFEST_FILE:-/dev/null}" ]]; then
     if [[ -f "$0.runfiles_manifest" ]]; then
@@ -20,7 +21,25 @@ else
   exit 1
 fi
 # --- end runfiles.bash initialization ---
-
-setup_truffle="$(rlocation __main__/src/blockchain/setup_truffle.sh)"
-source $setup_truffle
-$truffle compile
+adapter_server_args=()
+coordinator_server_args=()
+# Needs to be consistent between JS adapter server and coordinator server.
+adapter_server_port=51560
+while [ $# -gt 0 ]
+do
+  case "$1" in
+    --adapter_server_port*) adapter_server_port=${1#*=};;
+    --) shift; break;;
+    *) adapter_server_args+=" $1";;
+  esac
+  shift
+done
+while [ $# -gt 0 ]
+do
+  coordinator_server_args+=" $1"
+  shift
+done
+adapter_server_args+=" --adapter_server_port=$adapter_server_port"
+coordinator_server_args+=" --blockchain_adapter_port=$adapter_server_port"
+$(rlocation __main__/src/blockchain/client/run_adapter_server_command.bash) $adapter_server_args
+$(rlocation __main__/src/coordinator/coordinator_server_main) $coordinator_server_args

### 2022-06-01 src/blockchain/client/adapter_server.js
Add end to end script for testing client with 2 cohort transactions with the blockchain.

@@ -97,7 +97,7 @@ function main() {
   server.bindAsync(
       '0.0.0.0:' + port, grpc.ServerCredentials.createInsecure(), () => {
         server.start();
-        console.log('Started server');
+        console.log('Started blockchain adapter server');
       });
 }
 

### 2022-06-01 src/blockchain/client/run_adapter_server.sh

@@ -35,8 +35,6 @@ while [ $# -gt 0 ]
 do
   flag=${1%%=*}
   value="${1#*=}"
-  echo $flag
-  echo $value
   case "$flag" in
     --contract_address*) contract_address="$value";;
     --adapter_server_port*) adapter_server_port="$value";;

### 2022-06-01 src/blockchain/two_phase_commit.h

@@ -20,8 +20,12 @@ class TwoPhaseCommit {
  public:
   explicit TwoPhaseCommit(std::shared_ptr<grpc::Channel> channel)
       : TwoPhaseCommit(TwoPhaseCommitAdapter::NewStub(channel)) {
+    int failure_count = 0;
     while (!GetHeartBeat().ok()) {
-      LOG(INFO) << "Failed to get heartbeat. Waiting another second.";
+      ++failure_count;
+      if (failure_count % 10 == 0) {
+        LOG(INFO) << "Still waiting for blockchain adapter server";
+      }
       std::this_thread::sleep_for(absl::ToChronoSeconds(absl::Seconds(1)));
     }
   }

### 2022-06-01 src/client/client.h

@@ -26,6 +26,8 @@ class Client {
   ResponseOrStatus CommitSync(
       const coordinator::CommitAtomicTransactionRequest& request);
 
+  void WaitForCoordinator();
+
  private:
   bool TryGetTransactionResults(
       const coordinator::GetTransactionResultRequest& request,

### 2022-06-01 src/client/client_main.cc

@@ -5,6 +5,7 @@
 
 #include "absl/container/flat_hash_map.h"
 #include "absl/container/flat_hash_set.h"
+#include "absl/flags/flag.h"
 #include "absl/status/statusor.h"
 #include "absl/strings/escaping.h"
 #include "absl/strings/str_cat.h"
@@ -18,9 +19,11 @@
 #include "src/client/client.pb.h"
 #include "src/proto/common.pb.h"
 #include "src/proto/coordinator.grpc.pb.h"
-#include "tools/cpp/runfiles/runfiles.h"
 
-constexpr char* kCoordinatorPort = "58000";
+ABSL_FLAG(std::string, coordinator_port, "58000",
+          "Port to send requests to the coordinator");
+ABSL_FLAG(std::string, request_file_prefix, "two_namespaces_simple",
+          "Prefix of filename with requests to send to the coordinator");
 
 struct RunInfo {
   absl::flat_hash_set<std::string> cohort_addresses;
@@ -194,39 +197,14 @@ void LogResponseOrStatus(const client::ResponseOrStatus& response_or_status) {
   }
 }
 
-void InitializeServers(
-    const absl::flat_hash_set<std::string>& cohort_addresses) {
-  for (const std::string& address : cohort_addresses) {
-    if (address.substr(0, 8) == "0.0.0.0:") {
-      const std::string port = address.substr(8);
-      pid_t pid = fork();
-      if (pid == 0) {
-        // TODO(benjmarks22): Run the start_cohort_with_blockchain_adapter
-        // script. cohort::RunServer(port, /*num_db_threads=*/2,
-        //                   absl::StrCat("/tmp/data/", port),
-        //                   absl::StrCat("/tmp/txn_responses/", port));
-        while (true) {
-        }
-      }
-    }
-  }
-  pid_t pid = fork();
-  if (pid == 0) {
-    // TODO(benjmarks22): Run the start_coordinator_with_blockchain_adapter
-    // script. coordinator::RunServer(
-    //     kCoordinatorPort,
-    //     /*default_presumed_abort_duration=*/absl::Minutes(1));
-    while (true) {
-    }
-  }
-}
-
 void SendRequests(
+    const std::string& coordinator_port,
     std::vector<std::pair<coordinator::CommitAtomicTransactionRequest, bool>>&
         requests) {
   client::Client client(
-      grpc::CreateChannel(absl::StrCat("0.0.0.0:", kCoordinatorPort),
[…]

### 2022-06-01 src/client/end_to_end_test.sh

@@ -1,0 +1,91 @@
+#!/bin/bash
+set -eEuo pipefail
+
+# --- begin runfiles.bash initialization ---
+if [[ ! -d "${RUNFILES_DIR:-/dev/null}" && ! -f "${RUNFILES_MANIFEST_FILE:-/dev/null}" ]]; then
+    if [[ -f "$0.runfiles_manifest" ]]; then
+      export RUNFILES_MANIFEST_FILE="$0.runfiles_manifest"
+    elif [[ -f "$0.runfiles/MANIFEST" ]]; then
+      export RUNFILES_MANIFEST_FILE="$0.runfiles/MANIFEST"
+    elif [[ -f "$0.runfiles/bazel_tools/tools/bash/runfiles/runfiles.bash" ]]; then
+      export RUNFILES_DIR="$0.runfiles"
+    fi
+fi
+if [[ -f "${RUNFILES_DIR:-/dev/null}/bazel_tools/tools/bash/runfiles/runfiles.bash" ]]; then
+  source "${RUNFILES_DIR}/bazel_tools/tools/bash/runfiles/runfiles.bash"
+elif [[ -f "${RUNFILES_MANIFEST_FILE:-/dev/null}" ]]; then
+  source "$(grep -m1 "^bazel_tools/tools/bash/runfiles/runfiles.bash " \
+            "$RUNFILES_MANIFEST_FILE" | cut -d ' ' -f 2-)"
+else
+  echo >&2 "ERROR: cannot find @bazel_tools//tools/bash/runfiles:runfiles.bash"
+  exit 1
+fi
+# --- end runfiles.bash initialization ---
+
+function start_blockchain_with_ganache() {
+    rm /tmp/ganache_log.txt
+    deploy_ganache="$(rlocation __main__/src/blockchain/deploy_ganache_command.bash)"
+    $deploy_ganache $1 $2 $3 > /tmp/ganache.txt &
+    while [ 1 ]
+    do
+        contract_address=$(grep "Contract created:" /tmp/ganache_log.txt | tail -n 1 | cut -d " " -f 5) || :
+        if [ -n "$contract_address" ]
+        then
+            echo "$contract_address"
+            return
+        fi
+        sleep 2
+    done
+}
+
+function start_cohort() {
+    cohort_port=$3
+    adapter_server_port=$((cohort_port+500))
+    fuser -k $cohort_port/tcp && sleep 1s
[…]

### 2022-06-01 src/cohort/cohort_server_main.cc

@@ -1,3 +1,4 @@
+#include <filesystem>
 #include <string>
 #include <thread>
 
@@ -28,6 +29,8 @@ void RunServer(const std::string& port,
                const std::string& blockchain_adapter_port, uint num_db_threads,
                const std::string& db_data_dir,
                const std::string& db_txn_response_dir) {
+  std::filesystem::create_directories(db_data_dir);
+  std::filesystem::create_directories(db_txn_response_dir);
   std::string server_address = absl::StrCat("0.0.0.0:", port);
   std::string blockchain_adapter_address =
       absl::StrCat("0.0.0.0:", blockchain_adapter_port);

### 2022-06-01 src/cohort/start_cohort_with_blockchain_adapter_server.sh

@@ -43,4 +43,4 @@ done
 adapter_server_args+=" --adapter_server_port=$adapter_server_port"
 cohort_server_args+=" --blockchain_adapter_port=$adapter_server_port"
 $(rlocation __main__/src/blockchain/client/run_adapter_server_command.bash) $adapter_server_args
-$(rlocation __main__/src/cohort/cohort_server_main) $cohort_server_args
+$(rlocation __main__/src/cohort/cohort_server_main) $cohort_server_args &

### 2022-06-01 src/coordinator/coordinator_server.cc

@@ -172,10 +172,6 @@ void CoordinatorServer::SendCohortPrepareRequests(
     metadata.single_cohort_namespace = sub_transactions[0].namespace_;
   }
   size_t cohort_index = 0;
-  std::vector<std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-      cohort::PrepareTransactionResponse>>>
-      rpcs;
-  rpcs.reserve(sub_transactions.size());
   for (const SubTransaction &sub_transaction : sub_transactions) {
     if (sub_transactions.size() != 1) {
       prepare_request.set_cohort_index(cohort_index);
@@ -221,36 +217,6 @@ void CoordinatorServer::PrepareCohortTransaction(
       .PrepareTransaction(
           metadata_by_transaction_[transaction_id].contexts.back().get(),
           request, &response);
-}
-
-grpc::Status CoordinatorServer::FinishPrepareCohortTransaction(
-    std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-        cohort::PrepareTransactionResponse>> &async_response,
-    cohort::PrepareTransactionResponse &response) {
-  grpc::Status status;
-  async_response->Finish(&response, &status,
-                         /*tag=*/static_cast<void *>(&async_response));
-  return status;
-}
-
-std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-    cohort::GetTransactionResultResponse>>
-CoordinatorServer::AsyncGetResultsFromCohort(
-    const Namespace &namespace_,
-    const cohort::GetTransactionResultRequest &request,
-    ClientContext &context) {
-  return GetCohortStub(namespace_)
-      .AsyncGetTransactionResult(&context, request, /*cq=*/nullptr);
-}
-
-grpc::Status CoordinatorServer::FinishAsyncGetResultsFromCohort(
-    std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-        cohort::GetTransactionResultResponse>> &async_response,
-    cohort::GetTransactionResultResponse &response) {
-  grpc::Status status;
-  async_response->Finish(&response, &status,
-                         /*tag=*/static_cast<void *>(&async_response));
-  return status;
 }
 
 grpc::Status CoordinatorServer::GetResultsFromCohort(
@@ -322,32 +288,20 @@ void CoordinatorServer::UpdateCommittedResponseFromCohorts(
       response_by_transaction_[transaction_id].response;
[…]

### 2022-06-01 src/coordinator/coordinator_server.h

@@ -65,19 +65,6 @@ class CoordinatorServer : public Coordinator::Service {
       const std::string &transaction_id, const common::Namespace &namespace_,
       const cohort::PrepareTransactionRequest &request,
       const grpc::ServerContext &context);
-  virtual grpc::Status FinishPrepareCohortTransaction(
-      std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-          cohort::PrepareTransactionResponse>> &async_response,
-      cohort::PrepareTransactionResponse &response);
-  virtual std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-      cohort::GetTransactionResultResponse>>
-  AsyncGetResultsFromCohort(const common::Namespace &namespace_,
-                            const cohort::GetTransactionResultRequest &request,
-                            grpc::ClientContext &context);
-  virtual grpc::Status FinishAsyncGetResultsFromCohort(
-      std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-          cohort::GetTransactionResultResponse>> &async_response,
-      cohort::GetTransactionResultResponse &response);
   virtual grpc::Status GetResultsFromCohort(
       const common::Namespace &namespace_,
       const cohort::GetTransactionResultRequest &request,

### 2022-06-01 src/coordinator/coordinator_server_test.cc

@@ -35,8 +35,6 @@ class CoordinatorWithMockCohorts : public coordinator::CoordinatorServer {
                     const common::Namespace& namespace_,
                     const cohort::PrepareTransactionRequest& request,
                     const grpc::ServerContext& context));
-  MOCK_METHOD1(MockFinishAsyncGetResultsFromCohort,
-               grpc::Status(cohort::GetTransactionResultResponse& response));
   MOCK_METHOD4(MockGetResultsFromCohort,
                grpc::Status(const common::Namespace& namespace_,
                             const cohort::GetTransactionResultRequest& request,
@@ -60,20 +58,6 @@ class CoordinatorWithMockCohorts : public coordinator::CoordinatorServer {
       const cohort::PrepareTransactionRequest& request,
       const grpc::ServerContext& context) override {
     MockPrepareCohortTransaction(transaction_id, namespace_, request, context);
-  }
-  std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-      cohort::GetTransactionResultResponse>>
-  AsyncGetResultsFromCohort(
-      const common::Namespace& /*namespace_*/,
-      const cohort::GetTransactionResultRequest& /*request*/,
-      grpc::ClientContext& /*context*/) override {
-    return nullptr;
-  }
-  grpc::Status FinishAsyncGetResultsFromCohort(
-      std::unique_ptr<grpc::ClientAsyncResponseReaderInterface<
-          cohort::GetTransactionResultResponse>>& /*async_response*/,
-      cohort::GetTransactionResultResponse& response) override {
-    return MockFinishAsyncGetResultsFromCohort(response);
   }
   grpc::Status GetResultsFromCohort(
       const common::Namespace& namespace_,
@@ -377,10 +361,10 @@ TEST(CoordinatorServerTest, GetsResultsWhenCommitted) {
   cohort_get_response->mutable_get()->set_key("b");
   cohort_get_response->mutable_namespace_()->set_address("namespace2");
   cohort_get_response->mutable_value()->set_int64_value(3);
-  EXPECT_CALL(server, MockFinishAsyncGetResultsFromCohort(_))
-      .WillOnce(DoAll(SetArgReferee<0>(mock_cohort_response),
+  EXPECT_CALL(server, MockGetResultsFromCohort(_, _, _, _))
[…]

### 2022-06-01 src/coordinator/start_coordinator_with_blockchain_adapter_server.sh

@@ -42,4 +42,4 @@ done
 adapter_server_args+=" --adapter_server_port=$adapter_server_port"
 coordinator_server_args+=" --blockchain_adapter_port=$adapter_server_port"
 $(rlocation __main__/src/blockchain/client/run_adapter_server_command.bash) $adapter_server_args
-$(rlocation __main__/src/coordinator/coordinator_server_main) $coordinator_server_args
+$(rlocation __main__/src/coordinator/coordinator_server_main) $coordinator_server_args &

### 2022-06-01 paper/paper.tex
Add paper draft tex and pdf.

@@
-\documentclass{article}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\begin{document}
 
-\begin{document}
-Hello! Today!
+\title{Two-Phase Commit Using Blockchain}
+
+\author{\IEEEauthorblockN{Benjamin Jacob Marks}
+\IEEEauthorblockA{\textit{Stanford University} \\
+benmarks@stanford.edu}
+\and
+\IEEEauthorblockN{Heron Yang}
+\IEEEauthorblockA{\textit{Stanford University} \\
+heronyang@stanford.edu}
+\and
+\IEEEauthorblockN{Yuewei Na}
+\IEEEauthorblockA{\textit{Stanford University} \\
+yueweina@stanford.edu}
+}
+
+\maketitle
+
+\begin{abstract}
+We created a decentralized system that uses a public blockchain as a two-phase commit (2PC) coordinator, so users can securely commit an atomic transaction across any databases that support our interface.
+Users treat our system as one large database with ACID, despite being composed of many individual databases that are unaware of each other’s existence.
+By using a blockchain as a coordinator, it is resilient to failures such as network outages, power outages, and disk failures that cause traditional 2PC implementations to block.
+Our initial integration is with the Ethereum blockchain and LMDB databases; however, our system provides a generic interface to integrate with other blockchains and databases.
+\end{abstract}
+
+\begin{IEEEkeywords}
+atomic commit, blockchain, smart contract, distributed system, database transaction
+\end{IEEEkeywords}
+
+\section{Introduction}
+Two-Phase Commit (2PC) is a widely-adopted protocol that allows distributed systems to reach the consensus of distributed transactions.
+It ensures that the transaction can be either committed or aborted at all distributed machines.
+However, blocking is proven [3] to be to be inevitable even in
+synchronous systems where bounds on delays can be reliably estimated.
+
+Since the invention of Bitcoin, the underlying technology, the blockchain, not only gained attraction from cryptocurrency systems like Ethereum but also academic research about computer systems.
+At the core of blockchain, it is a protocol of reaching consensus by using Proof-of-Work or Proof-of-Stake.
+Previous work [1] has shown that synchronous blockchain can only increase the reliability of the original 2PC protocol but also make it non-blocking when various kinds of failures happen.
+
+In this paper, we further extended the previous work with an optimized system architecture and properly designed interfaces.
+These contributions can make 2PC with blockchain work in production and at scale.
+
+\subsection{Key Client Features}
+
+\subsubsection{Namespaced Distributed Transaction}
+We shard the data space with namespaces.
+Each cohort is only responsible for one namespace.
+A client transaction is composed of a series of <namespace, operation> pairs.
+An operation is either Put(key, value) or Get(key).
+The Coordinator will forward the operation to the cohort(s) that’s responsible for that namespace.
+Our system guarantees that the transaction can be committed or aborted with consensus from multiple distributed cohorts.
+
+\subsubsection{Simple Async Interface}
+To a client, they only need to first send CommitAtomicTransactionRequest to the coordinator and then send request GetTransactionResult to fetch the result of their transaction.
+
+\subsection{System Properties}
+Compared to the traditional version of 2PC protocol, our approach increases extensibility, scalability and reliability.
+
+\subsubsection{Extensibility}
+We can plug any blockchain, regardless of being public permissionless or private permissioned into our blockchain interface - `StartVoting`, `Vote`, `GetVotingDecision`.
+We can also plug any database system that has ACID support into our transaction delegation interface.
+The transaction delegation interface implements most common transaction operations - `Begin`, `Commit`, `Abort`, `Get`, `Put`.
+
+\subsubsection{Scalability}
+Our Coordinator is designed to be light-weight to handle high throughput.
+This is achieved by delegating 2 parts of work in traditional 2PC [2].
+First, the coordination and voting work is done in the blockchain.
+Second, it divides the client transaction into shards and forwards to the corresponding cohort that has the data.
+
+Adding a cohort to the system is as simple as making the coordinator aware of the address of the newly added cohort. The system is horizontally scalable.
+
+\subsubsection{Reliability}
+With the introduction of namespace, it can improve cohort reliability by adding a cohort whose namespace is the same with one or more of the existing cohorts.
+This essentially adds another replication in the state replication machine.
+
+On the other hand, leveraging permissionless blockchain increases coordinator reliability. When a coordinator crashes, the commitment states are all kept in the blockchain.
+
+\subsection{Design Overview}
+3.1 Architecture and Components
+We define a transaction as a list of operations starting with BEGIN() followed by multiple GET() or PUT() and ending with COMMIT().
+Client creates a transaction and requests Coordinator to commit.
+Coordinator distributes the transaction to corresponding Cohorts to prepare then offload the remaining 2PC responsibility to the blockchain.
+Each cohort prepares the transaction locally, votes to Blockchain if ready, checks on Blockchain to see if a transaction should be committed, then finally commits the transaction to its database.
+
+The architecture is similar to the traditional 2PC architecture where we have one coordinator and multiple cohorts - except each cohort talks to the blockchain directly to vote and to look up the voting decision as the coordinator offload the 2PC voting responsibility to the blockchain.
+
+We describe details of each component and each in the remaining section 4.1, the complete flow of handling a successful transaction in Section 4.2, and how we handle failures in Section 4.3.
+
+Fig.
+Client starts a transaction request with Coordinator.
+Coordinator distributes the transaction to Cohorts and offload the 2PC responsibility to blockchain.
+Cohort interacts with the blockchain directly to vote and to look up a transaction status.
+
+\subsubsection{Blockchain}
+The blockchain component performs a 2PC protocol so it’s responsible for tracking voting states of each transaction and making decisions of committing.
+The implementation uses Ethereum smart contract to store the followings states of each transaction (indexed by the transaction ID) internally:
+$transaction_state$ Voting states of each cohort of a transaction.
+$transaction_config$ Transaction configure of a transaction including the number of cohorts and the timeout time.
+
+And, the smart contract exposes the following APIs externally:
+$StartVoting(transaction_id, cohorts, vote_timeout_time)$ Called by the coordinator to start voting on a transaction.
+`cohorts` is the number of cohorts that will vote on this transaction later.
+$vote_timeout_time$ is the absolute time when the 2PC stops taking.
+$Vote(transaction_id, cohort_id, ballot)$ Called by the cohort to vote COMMIT or ABORT on a transaction before timeout.
+$GetVotingDecision(transaction_id)$ Called by any component to check the voting state on a transaction.
+
+StartVoting() initializes the states for a transaction and Vote() updates the voting status of a transaction. GetVotingDecision() makes the decision by following the below logic:
+If any ABORT vote is received, return ABORT decision.
+If all COMMIT votes are received, return COMMIT decision.
+If not (1) and not (2), returns ABORT decision if timed out; otherwise, return PENDING decision.
+
+\subsubsection{Database}
+// heron: should introduce database key, namespace, lock.
+
+\subsubsection{Coordinator}
+// heron: should describe sub-transaction.
+
+\subsubsection{Cohort}
+
+\subsection{System Interactions (a.k.a Life of a Transaction)}
+This section describes the end-to-end life of a transaction without handling failures.
+
+\begin{enumerate}
+\item Client composes a transaction request and passes it to Coordinator through `CommitAomticTransaction()`.
+Coordinator returns a transaction ID for the client to look up the transaction result later.
+\item Coordinator does the followings in parallel: (a) Coordinator decomposes the transaction into sub-transaction and distributes them to the corresponding cohort through `PrepareTransaction()`.
+(b) Coordinator asks Blockchain to trace votes for the new transaction through `StartVoting()`.
+\item Cohort executes a sub-transaction by forwarding the operations to the database up to the point before Commit(). Cohort writes an entry to both the undo and redo log.
+\item Cohort votes to Blockchain through `Vote()` if it wants to commit or abort the transaction.
+\item Cohort sleeps until the expected transaction timeout time and gets the voting decision from Blockchain through `GetVotingDecision()`.
+\item If the voting decision is to commit the transaction, cohort commits the transaction onto the database and releases acquired locks; otherwise, undoes the transaction using the undo log and releases acquired locks.
+If a transaction is committed, the cohort writes the GET operation results to disk.
+\item (Independent to the previous steps) Client has waited until the timeout time and requests the transaction result from Coordinator through `GetTransactionResult()`, using the transaction ID given in step 1.
+\item Coordinator gets (1) the voting decision of the inquiry transaction from Blockchain through `GetTransactionResult()` and (2) the results from the disk written by the cohort.
+\end{enumerate}
+
+\subsection{Handling Failures}
+
+// How does the whole system deal with crash failures?
+\subsubsection{Crash of coordinator}
+\subsubsection{Crash of cohort}
+\subsubsection{Data loss of database}
+\subsubsection{Slow blockchain}
+We use Ethereum for our blockchain implementation.
+According to [this article], it takes around 15 seconds to 5 minutes to process a blockchain transaction on Ethereum, which indicates that the blockchain is relatively slow compared to other components within our system.
+
+A slow blockchain affects the number of COMMIT votes to collect before the timeout time, which leads to more transactions to abort eventually; however, the blockchain slowness doesn’t affect the promised atomicity.
+The client or the coordinator can specify a longer timeout time if needed.
+
+\section{Implementation}
+
+\subsection{Environment and Tools}
+// Which tool, machine, etc.
+
+\subsubsection{Blockchain}
+We wrote our two-phase-commit smart contract with Solidity and developed it with Truffle [https://trufflesuite.com/].
+With Truffle, we compile our smart contract and deploy it onto a local blockchain simulator.
+
+To interact with the deploy smart contract, we need a smart contract client and be used by the coordinator and cohorts.
+Since we implement most of the code in C++ and failed to find a C++ smart contract client library, we use a JavaScript library (web3.js [https://web3js.readthedocs.io/en/v1.7.3/]) instead and add an additional RPC layer for our C++ code to use.
+
+\subsection{Testing and Evaluation}
+// How do we know if the system is performing correctly?
+
+\subsubsection{Blockchain Cost}
+“Gas Used” is the fee the blockchain caller pays to the miner for processing the transaction.
+In the Truffle development environment, we simulate the scenario where we have two cohorts voting to commit, the total gas we pay is around 0.33 USD [reference: https://automatedwebtools.com/gwei-to-usd-calculator/].
+
+\section{Discussion}
+// What have we done? What went well and what didn’t?
+// Future directions?
+
 \end{document}
+\endinput

### 2022-06-01 setup.sh

@@ -25,7 +25,7 @@ function install () {
 
 BINARY_DIR="/usr/local/bin"
 $SUDO apt-get update || exit 2
-$SUDO apt-get install -y clang-format clang-tidy curl git gcc g++ python3 python3-pip npm texlive || exit 3
+$SUDO apt-get install -y clang-format clang-tidy curl git gcc g++ python3 python3-pip npm texlive texlive-generic-extra texlive-plain-generic texlive-science || exit 3
 install ${BINARY_DIR} bazel "https://github.com/bazelbuild/bazelisk/releases/download/v1.11.0/bazelisk-linux-amd64"
 install ${BINARY_DIR} buildifier "https://github.com/bazelbuild/buildtools/releases/download/5.1.0/buildifier-linux-amd64"
 $SUDO npm install -g prettier

### 2022-06-03 paper/paper.tex
Update the font to 12pt.
Update the font to 11pt.

--- 2022-05-30 paper/paper.tex
@@
-\documentclass{article}
+\documentclass[11pt,conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\begin{document}
 
-\begin{document}
-Hello! Today!
+\title{Two-Phase Commit Using Blockchain}
+
+\author{\IEEEauthorblockN{Benjamin Jacob Marks}
+\IEEEauthorblockA{\textit{Stanford University} \\
+benmarks@stanford.edu}
+\and
+\IEEEauthorblockN{Heron Yang}
+\IEEEauthorblockA{\textit{Stanford University} \\
+heronyang@stanford.edu}
+\and
+\IEEEauthorblockN{Yuewei Na}
+\IEEEauthorblockA{\textit{Stanford University} \\
+yueweina@stanford.edu}
+}
+
+\maketitle
+
+\begin{abstract}
+We created a decentralized system that uses a public blockchain as a two-phase commit (2PC) coordinator, so users can securely commit an atomic transaction across any databases that support our interface.
+Users treat our system as one large database with ACID, despite being composed of many individual databases that are unaware of each other’s existence.
+By using a blockchain as a coordinator, it is resilient to failures such as network outages, power outages, and disk failures that cause traditional 2PC implementations to block.
+Our initial integration is with the Ethereum blockchain and LMDB databases; however, our system provides a generic interface to integrate with other blockchains and databases.
+\end{abstract}
+
+\begin{IEEEkeywords}
+atomic commit, blockchain, smart contract, distributed system, database transaction
+\end{IEEEkeywords}
+
+\section{Introduction}
+Two-Phase Commit (2PC) is a widely-adopted protocol that allows distributed systems to reach the consensus of distributed transactions.
+It ensures that the transaction can be either committed or aborted at all distributed machines.
+However, blocking is proven [3] to be to be inevitable even in
+synchronous systems where bounds on delays can be reliably estimated.
+
+Since the invention of Bitcoin, the underlying technology, the blockchain, not only gained attraction from  cryptocurrency systems like Ethereum but also academic research about computer systems.
+At the core of blockchain, it is a protocol of reaching consensus by using Proof-of-Work or Proof-of-Stake.
+Previous work [1] has shown that synchronous blockchain can only increase the reliability of the original 2PC protocol but also make it non-blocking when various kinds of failures happen.
+
+In this paper, we further extended the previous work with an optimized system architecture and properly designed interfaces.
+These contributions can make 2PC with blockchain work in production and at scale.
+Compared to the traditional version of 2PC protocol [2], our proposed approach increases extensibility, scalability and reliability.
+
+\subsubsection{Extensibility}
+We can plug any blockchain, regardless of being public permissionless or private permissioned into our blockchain interface - \texttt{StartVoting}, \texttt{Vote}, \texttt{GetVotingDecision}.
+We can also plug any database system that has ACID support into our transaction delegation interface.
+The transaction delegation interface implements most common transaction operations - \texttt{Begin}, \texttt{Commit}, \texttt{Abort}, \texttt{Get}, \texttt{Put}.
+
+\subsubsection{Scalability}
+Our Coordinator is designed to be light-weight to handle high throughput and horizontally scalable.
+This is achieved by delegating coordination and voting work to the blockchain.
+The delegation makes the coordinator stateless and allows many coordinator jobs to co-exist to balance the load and tolerate job failures.
+To cohort, adding a cohort to the system is as simple as making the coordinator be aware of the address of the newly added cohort.
+So our system is horizontally scalable.
+
+\subsubsection{Reliability}
+Our system has significantly better reliability than the traditional 2PC protocol for three reasons.
+First, the system leverages blockchain to maintain the coordination state.
+The stateless coordinator can easily recover as long as it knows the participating cohorts.
+Second, The coordinator is stateless and horizontally scalable.
+We can keep multiple jobs running concurrently and place a load-balancing layer on top of it.
+Third, with the introduction of $namespace$ (see \ref{coordinator}), it can improve cohort reliability by adding a cohort whose namespace is the same with one or more of the existing cohorts.
+This essentially adds another replication in the state replication machine.
+
+\section{Design Overview}
+\subsection{Architecture and Components} \label{components}
+We define a transaction as a list of operations starting with \texttt{BEGIN} followed by multiple \texttt{GET} or \texttt{PUT} and ending with \texttt{COMMIT}.
+Client creates a transaction and requests Coordinator to commit.
+Coordinator distributes the transaction to corresponding Cohorts to prepare then offload the remaining 2PC responsibility to the blockchain.
+Each cohort prepares the transaction locally, votes to Blockchain if ready, checks on Blockchain to see if a transaction should be committed, then finally commits the transaction to its database.
+
+The architecture is similar to the traditional 2PC architecture where we have one coordinator and multiple cohorts - except each cohort talks to the blockchain directly to vote and to look up the voting decision as the coordinator offload the 2PC voting responsibility to the blockchain.
+
+We describe details of each component and each in the remaining \ref{components}, the complete flow of handling a successful transaction in \ref{txnflow}, and how we handle failures in \ref{failurecases}.
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.4\textwidth]{./diagram.png}
+\caption{Client starts a transaction request with Coordinator.
+Coordinator distributes the transaction to Cohorts and offload the 2PC responsibility to blockchain.
+Cohort interacts with the blockchain directly to vote and to look up a transaction status.}
+\label{fig:Panel}
+\end{figure}
+
+\subsubsection{Blockchain} \label{blockchain}
+The blockchain component performs a 2PC protocol so it’s responsible for tracking voting states of each transaction and making decisions of committing.
+The implementation uses Ethereum smart contract to store the followings states of each transaction (indexed by the transaction ID) internally:
+$transaction\_state$ Voting states of each cohort of a transaction.
+$transaction\_config$ Transaction configure of a transaction including the number of cohorts and the timeout time.
+
+We designed a \texttt{\\BlockchainAdapter} interface. It exposes the following APIs externally:
+\texttt{StartVoting(\\transaction\_id, cohorts, vote\_timeout\_time)} Called by the coordinator to start voting on a transaction.
+$cohorts$ is the number of cohorts that will vote on this transaction later.
+$vote\_timeout\_time$ is the absolute time when the 2PC stops taking.
+\texttt{Vote(transaction\_id, cohort\_id, ballot)} Called by the cohort to vote $COMMIT$ or $ABORT$ on a transaction before timeout.
+\texttt{GetVotingDecision(transaction\_id)} Called by any component to check the voting state on a transaction.
+
+\texttt{StartVoting} initializes the states for a transaction and \texttt{Vote} updates the voting status of a transaction. \texttt{GetVotingDecision} makes the decision by following the below logic:
+If any $ABORT$ vote is received, return $ABORT$ decision.
+If all $COMMIT$ votes are received, return $COMMIT$ decision.
+If not (1) and not (2), returns $ABORT$ decision if timed out; otherwise, return $PENDING$ decision.
+
+\subsubsection{Database} \label{database}
+The database component is responsible for committing sub-transactions in each cohort. It abstracts typical ACID databases with 5 APIs in a DatabaseAdapter 
+\begin{itemize}
+  \item \texttt{Begin()}:  starts a transaction
+  \item \texttt{Commit()}: commits a started transaction
+  \item \texttt{Abort()}: aborts a started transaction
+  \item \texttt{Put(key, value)}: puts value in a row whose primary key is key 
+  \item \texttt{Get(key)}: gets the value from the row whose primary key is key 
+\end{itemize}
+
+We require the underlying database to be ACID because the database can handle locks and transactions natively without making cohorts implementing complicated locking logic for Put and Get operations.
+But for databases that only support weak snapshot isolation (e.g.
+LMDB [4]), the cohort only needs a single write lock to prevent other write transactions to be started to make sure that transactions are serialized.
+
+\subsubsection{Coordinator} \label{coordinator}
+The Coordinator exposes two APIs to the clients
+\begin{itemize}
+  \item \texttt{CommitAtomicTransaction(
+    CommitAtomicTransactionRequest, CommitAtomicTransactionResponse)}
+  \item \texttt{GetTransactionResult(
+    GetTransactionResultRequest, GetTransactionResultResponse)}
+\end{itemize}
+
+When a coordinator receives a \texttt{\\CommitAtomicTransaction} call from the client, it will first decompose the transaction into sub-transactions, which will be executed in the cohorts, based on the $namespace$ information.
+We shard the data key space into $namespaces$.
+Each cohort is responsible for a range of keys under the same $namespace$.
+After transaction decomposition, it will send \texttt{PrepareTransaction} requests with the sub-transaction info to the cohorts.
+At the same time, it will send \texttt{StartVoting} to the blockchain to trace the voting status from the cohorts.
+Different from the traditional 2PC protocol, our coordinator is designed to delegate voting and coordination work to the blockchain.
+From here on, it no longer handles any work unless \texttt{GetTransactionResult} is requested from the client.
+
+When the client sends \texttt{GetTransactionResult} to the coordinator, it reads the transaction status from the blockchain.
+If the status is $PENDING$ or $ABORTED$, it will return the status directly.
+If the transaction is $COMMITTED$, it will further send requests to the cohorts to fetch the Get result as well and return the status and the get results altogether.
+If not all cohorts return the Get result (e.g.
+due to crashes or network latency), the coordinator will send a partial response to the client with the ones that responded along with and indicator that the response is not yet complete.
+
+\subsubsection{Cohort} \label{cohort}
+
+The Cohort exposes two APIs to the Coordinator
+\begin{itemize}
+  \item \texttt{PrepareTransaction(
+    PrepareTransactionRequest, PrepareTransactionResponse)}
+  \item \texttt{GetTransactionResult(
+    GetTransactionResultRequest, GetTransactionResultResponse)}
+\end{itemize}
+
+The $PrepareTransactionRequest$ contains a $transaction\_id$, a list of operations for the sub-transaction and an abort deadline.
+When a cohort receives a PrepareTransaction request from the coordinator, it takes these steps
+\begin{enumerate}
+  \item Start a new thread in the thread pool.
+  \item Creates a \texttt{DatabaseAdapter} object.
+  \item Starts a transaction with 
+  \texttt{DatabaseAdapter\\::Begin()}. 
+  \item Do a series of requested Put or Get operations in the sub-transaction. 
+If any operation is reported invalid by the database, the cohort will send an $ABORT$ ballot to the blockchain via \texttt{BlockchainAdapter::Vote()}.
+  \item If all operations succeed, it will store the got values in both an in-memory hashmap as well as persistently on disk keyed by the $transaction\_id$ to prepare for future \texttt{GetTransactionResult}  calls.
+  \item Set a timer based on the input abort deadline.
+  \item If the timer reaches the deadline, it will send a request \texttt{BlockchainAdapter::\\GetVotingDecision(transaction\_id)} to the blockchain. 
+  \begin{enumerate}
+    \item If the blockchain states that the transaction should be committed, it will call \texttt{DatabaseAdapter::Commit()}.
+    \item If the blockchain states that the transaction should be aborted, it will call \texttt{DatabaseAdapter::Abort()}.
+  \end{enumerate}
+
+\end{enumerate}
+
+When the cohort receives a \texttt{GetTransaction\-Result} request, if it has committed the transaction, it simply reads the in-memory hashmap based on the $transaction\_id$ in the request and returns the got values as response.
+If it has not yet received a commit or abort response from the blockchain, it returns a pending response instead.
+
+\subsection{System Interactions (a.k.a Life of a Transaction)} \label{txnflow}
+This section describes the end-to-end life of a transaction without handling failures.
+
+\begin{enumerate}
+\item Client composes a transaction request and passes it to Coordinator through \texttt{\\CommitAomticTransaction}.
+Coordinator returns a transaction ID for the client to look up the transaction result later.
+\item Coordinator does the followings in parallel: (a) Coordinator decomposes the transaction into sub-transaction and distributes them to the corresponding cohort through \texttt{PrepareTransaction}.
+(b) Coordinator asks Blockchain to trace votes for the new transaction through \texttt{StartVoting}.
+\item Cohort executes a sub-transaction by forwarding the operations to the database up to the point before \texttt{Commit}. Cohort writes an entry to both the undo and redo log.
+\item Cohort votes to Blockchain through \texttt{Vote} if it wants to commit or abort the transaction.
+\item Cohort sleeps until the expected transaction timeout time and gets the voting decision from Blockchain through \texttt{GetVotingDecision}.
+\item If the voting decision is to commit the transaction, cohort commits the transaction onto the database and releases acquired locks; otherwise, undoes the transaction using the undo log and releases acquired locks.
+If a transaction is committed, the cohort writes the \texttt{GET} operation results to disk.
+\item (Independent to the previous steps) Client has waited until the timeout time and requests the transaction result from Coordinator through \texttt{GetTransactionResult}, using the transaction ID given in step 1.
+\item Coordinator gets (1) the voting decision of the inquiry transaction from Blockchain through \texttt{GetTransactionResult} and (2) the results from the memory or the disk written by the cohort.
+\end{enumerate}
+
+\subsection{Handling Failures} \label{failurecases}
+As described above, one of the main benefits of our system compared to traditional 2PC implementations is improved resiliency.
+By delegating the voting and coordination to the blockchain, our system will not block on any one machine that crashes or has a slow or partitioned network.
+
+\subsubsection{Crash of coordinator}
+The state of the coordinator includes only the list of which cohorts are involved in each active transaction.
+Thus if the coordinator crashes, when it recovers, it will respond to client requests for the Get responses of the transaction by indicating that it is unable to find the transaction.
+As a result, the client can directly ask the cohorts what the Get responses of the transaction are.
+Unlike in traditional 2PC, where the coordinator stores how each cohort has voted, in our design, this information is stored in the blockchain.
+Our system is resilient to such crashes, as the transaction does not need to block on the coordinator’s recovery.
+
+\subsubsection{Crash of cohort}
+Our system is also resilient to crashes of cohorts, as these crashes do not cause any other cohorts to block for longer than the $vote\_timeout\_time$.
+If a cohort crashes before sending a vote to commit, then the blockchain will wait until the $vote\_timeout\_time$ and all cohorts will abort the transaction.
+If a cohort crashes after voting to commit and the blockchain decides that the transaction should commit, then the cohort will replay all transactions that committed while it crashed and recovered.
+To replay the transactions, the cohort will acquire the same locks it had for the transactions and send the relevant put and get requests to the database and commit the transactions in the database.
+Since it has already computed the response, if the coordinator asks for the get responses for any of these transactions, the cohort can safely return them, even if they have not yet committed in the database.
+This can be implemented in the future but we decided not to include it in our first version.
+
+\subsubsection{Slow blockchain}
+We use Ethereum for our blockchain implementation.
+According to [this article], it takes around 15 seconds to 5 minutes to process a blockchain transaction on Ethereum, which indicates that the blockchain is relatively slow compared to other components within our system.
+
+A slow blockchain affects the number of COMMIT votes to collect before the timeout time, which leads to more transactions to abort eventually; however, the blockchain slowness doesn’t affect the promised atomicity.
+The client or the coordinator can specify a longer timeout time if needed.
+
+\section{Implementation}
+In this section we first describe the implementation of our overall system. Then we’ll illustrate the implementation of individual components.
+
+\subsection{Environment}
+We implement the coordinator and the cohorts with about 2800 lines of C++ code.
+For communications between components, we use gPRC and ProtoBuf.
+For providing the extensibility of swapping different database and blockchain implementations, we provide generic interfaces.
+We tested end-to-end behavior on a single machine with different processes simulating each component.
+
+\subsubsection{Blockchain}
+We wrote our two-phase-commit smart contract with Solidity and developed it with Truffle [\url{https://trufflesuite.com/}].
+Truffle provides the development environment for us to compile smart contracts, test it, and deploy it locally for debugging.
+Once we’ve deployed our smart contract, the coordinator and the cohorts can interact with the contract through a smart contract client library.
+However, we didn’t find a proper C++ smart contract client library to use, so we use a JavaScript library (web.js[\url{https://web3js.readthedocs.io/en/v1.7.3/}]) instead and add an additional gRPC layer to translate to C++.
+
+While testing in the Truffle development, we learned the gas used for successfully committing a transaction with two cohorts is around 0.33 USD [\href{https://automatedwebtools.com/gwei-to-usd-calculator/}{reference}] (see \ref{appendix} for breakdowns).
+
+\subsubsection{Database}
+We designed the DatabaseAdapter interface and implemented the methods with LMDB primitives.
+LMDB provides process-oriented functions like \texttt{open()}, \texttt{begin()}, \texttt{commit()}, \texttt{put()}, \texttt{get()}.
+Our object-oriented C++ implementation has almost 500 lines of code.
+We chose LMDB because it is fast, lightweight and easy to test in memory.
+But the interface is extensible to other databases like SQLite [6] as well.
+
+\subsubsection{Coordinator}
+The coordinator deterministically computes a global identifier for the transaction based on the client’s provided identifier and the client’s hostname and port.
+This allows the coordinator to avoid sending duplicate (potentially non-idempotent) requests to cohorts if the client sends the same request multiple times.
+This can happen if the client does not hear the response for its original request from the coordinator in time and tries to resend it.
+By using SHA256 to generate the transaction identifiers, it is extremely unlikely for transactions from different clients or with different client identifiers to have the same global identifier.
+
+\subsubsection{Cohort}
+One optimization we used for the cohort is that the \texttt{PrepareTransactionRequest} includes an indicator if this is the only cohort for the transaction.
+In such cases, we can avoid the expensive requests to the blockchain and the cohort can immediately commit a valid transaction as soon as it finishes preparing it in the database.
+This is safe because it knows that no other cohort can abort the transaction and that it is guaranteed to be atomic within its own database.
+
+\section{Discussion}
+
+\subsubsection{Latency tradeoffs}
+Our approach enables the client to tradeoff between latency and the likelihood of committing a transaction.
+When the client sets a longer time for timeout, it’s more likely to collect sufficient votes from all cohorts to make a COMMIT decision but may have higher latency if a cohort crashes.
+There is a limit to the minimum timeout, which can be no shorter than the time it takes to finalize a new blockchain block.
+
+As part of a future effort, we can extend the system to tradeoff between further latency reductions and safety by having cohorts send their votes to the coordinator as well as to the blockchain.
+If the coordinator receives commit votes from all cohorts, it can update the decision in the blockchain and tell all cohorts to commit.
+In this procedure, we would not need to wait for a new block to be mined in the blockchain to commit a transaction.
+The loss of safety comes from edge cases where the blockchain is slow to process the transaction and it ends up not receiving the votes in time and thus deciding to abort.
+If some cohorts received commit messages from the coordinator and others did not and then asked the blockchain for the decision, the cohorts who received commit messages would commit, while the others would abort the transaction.
+As we expect this to only affect a very small number of transactions, clients would be able to indicate if their transaction can tolerate a small chance of non-atomicity for much faster latency.
+
+\subsubsection{Distributed coordinator}
+Currently the coordinator is stateless, so it is easy to add new coordinator servers, as long as each one handles a disjoint set of transactions.
+However, if a coordinator crashes, then clients cannot ask other coordinators for the Get responses of a transaction, since those coordinators will not know how to find the cohorts for the transaction.
+As part of a future effort, we could have the coordinators send each other the cohorts for each transaction so that they can respond to client requests even if the original coordinator for that request crashes.
+
+\section{Reference}
+
+\section{Appendix} \label{appendix}
+\begin{table}[hbtp]
+     	\centering
+     	\caption{Gas used, cost in ETH/USD per blockchain transaction for a successful 2PC commit with 2 cohorts. Data as of 2022/05/30.}
+     	\label{table:cost}
+     	\begin{tabular}{|l|c|c|c|}
+     		\hline
+     		  & Gas Used (Gwei) &  Cost in ETH & Cost in USD \\
+     		\hline
+     		StartVoting &73441 &0.000073441 & 0.143\\
+     		\hline
+     		Vote 1 &56451	&0.000056451 &0.110\\
+     		\hline
+     		Vote 2 &39513 &0.000039513 &0.077\\
+     		\hline
+     	\end{tabular}
+     \end{table}
 \end{document}
+\endinput

### 2022-06-03 paper/paper.tex

@@
 Since the invention of Bitcoin, the underlying technology, the blockchain, not only gained attraction from  cryptocurrency systems like Ethereum but also academic research about computer systems.
-At the core of blockchain, it is a protocol of reaching consensus by using Proof-of-Work or Proof-of-Stake.
-Previous work [1] has shown that synchronous blockchain can only increase the reliability of the original 2PC protocol but also make it non-blocking when various kinds of failures happen.
+At its core, blockchain is a protocol of reaching consensus by using Proof-of-Work or Proof-of-Stake [//TODO: reference goes here].
+Previous work [1] has shown that synchronous blockchain can increase the reliability of the original 2PC protocol and also make it non-blocking when various kinds of failures happen.
 
 In this paper, we further extended the previous work with an optimized system architecture and properly designed interfaces.
-These contributions can make 2PC with blockchain work in production and at scale.
+These contributions make 2PC with blockchain work in production and at scale.
 Compared to the traditional version of 2PC protocol [2], our proposed approach increases extensibility, scalability and reliability.
@@
 \subsubsection{Extensibility}
-We can plug any blockchain, regardless of being public permissionless or private permissioned into our blockchain interface - \texttt{StartVoting}, \texttt{Vote}, \texttt{GetVotingDecision}.
-We can also plug any database system that has ACID support into our transaction delegation interface.
-The transaction delegation interface implements most common transaction operations - \texttt{Begin}, \texttt{Commit}, \texttt{Abort}, \texttt{Get}, \texttt{Put}.
+We can plug any blockchain, regardless of whether it is public/permissionless or private/permissioned, into our blockchain interface which exposes \texttt{StartVoting()}, \texttt{Vote()}, \texttt{GetVotingDecision()}.
+We can also plug any database system that has ACID support into our transaction delegation interface, which implements common transaction operations - \texttt{Begin()}, \texttt{Commit()}, \texttt{Abort()}, \texttt{Get()}, \texttt{Put()}.
 
 \subsubsection{Scalability}
-Our Coordinator is designed to be light-weight to handle high throughput and horizontally scalable.
+We design our coordinator  to be light-weight to handle high throughput and horizontally scalable.
 This is achieved by delegating coordination and voting work to the blockchain.
 The delegation makes the coordinator stateless and allows many coordinator jobs to co-exist to balance the load and tolerate job failures.
-To cohort, adding a cohort to the system is as simple as making the coordinator be aware of the address of the newly added cohort.
-So our system is horizontally scalable.
+Adding a cohort to the system is as simple as making potential clients aware of the address of the newly added cohort, so our system is horizontally scalable.
 
 \subsubsection{Reliability}
-Our system has significantly better reliability than the traditional 2PC protocol for three reasons.
+Our system has better reliability than the traditional 2PC protocol for two reasons.
 First, the system leverages blockchain to maintain the coordination state.
@@
 We can keep multiple jobs running concurrently and place a load-balancing layer on top of it.
-Third, with the introduction of $namespace$ (see \ref{coordinator}), it can improve cohort reliability by adding a cohort whose namespace is the same with one or more of the existing cohorts.
-This essentially adds another replication in the state replication machine.
 
@@
 We define a transaction as a list of operations starting with \texttt{BEGIN} followed by multiple \texttt{GET} or \texttt{PUT} and ending with \texttt{COMMIT}.
-Client creates a transaction and requests Coordinator to commit.
-Coordinator distributes the transaction to corresponding Cohorts to prepare then offload the remaining 2PC responsibility to the blockchain.
-Each cohort prepares the transaction locally, votes to Blockchain if ready, checks on Blockchain to see if a transaction should be committed, then finally commits the transaction to its database.
+The client creates a transaction and requests the coordinator to commit.
+The coordinator distributes the transaction to the corresponding cohorts to prepare and then offloads the remaining 2PC responsibility to the blockchain.
+Each cohort prepares the transaction locally, votes to the blockchain if ready, checks on Blockchain to see if a transaction should be committed, then finally commits the transaction to its database.
 
@@
 \includegraphics[width=0.4\textwidth]{./diagram.png}
-\caption{Client starts a transaction request with Coordinator.
-Coordinator distributes the transaction to Cohorts and offload the 2PC responsibility to blockchain.
-Cohort interacts with the blockchain directly to vote and to look up a transaction status.}
+\caption{Fig 1. This diagram displays the main components within the whole architecture and example steps of successfully committing a transaction.}
 \label{fig:Panel}
@@
 \subsubsection{Blockchain} \label{blockchain}
-The blockchain component performs a 2PC protocol so it’s responsible for tracking voting states of each transaction and making decisions of committing.
-The implementation uses Ethereum smart contract to store the followings states of each transaction (indexed by the transaction ID) internally:
-$transaction\_state$ Voting states of each cohort of a transaction.
-$transaction\_config$ Transaction configure of a transaction including the number of cohorts and the timeout time.
+The blockchain component performs a 2PC protocol, so it’s responsible for tracking voting states of each transaction and making decisions of committing.
+The implementation uses Ethereum smart contracts to store the followings states of each transaction (indexed by the transaction ID) internally:
+$transaction\_state$ stores voting states of each cohort of a transaction.
+$transaction\_config$ stores the transaction configure of each transaction including the number of cohorts and the vote timeout time.
 
-We designed a \texttt{\\BlockchainAdapter} interface. It exposes the following APIs externally:
+We designed a \texttt{\\BlockchainAdapter} interface, which exposes the following APIs externally to the coordinator and the cohort:
 \texttt{StartVoting(\\transaction\_id, cohorts, vote\_timeout\_time)} Called by the coordinator to start voting on a transaction.
 $cohorts$ is the number of cohorts that will vote on this transaction later.
-$vote\_timeout\_time$ is the absolute time when the 2PC stops taking.
+$vote\_timeout\_time$ is the absolute time on the blockchain when the 2PC stops taking votes.
 \texttt{Vote(transaction\_id, cohort\_id, ballot)} Called by the cohort to vote $COMMIT$ or $ABORT$ on a transaction before timeout.
@@
 \texttt{StartVoting} initializes the states for a transaction and \texttt{Vote} updates the voting status of a transaction. \texttt{GetVotingDecision} makes the decision by following the below logic:
-If any $ABORT$ vote is received, return $ABORT$ decision.
-If all $COMMIT$ votes are received, return $COMMIT$ decision.
-If not (1) and not (2), returns $ABORT$ decision if timed out; otherwise, return $PENDING$ decision.
+If any $ABORT$ vote is received, return an $ABORT$ decision.
+If all $COMMIT$ votes are received, return a $COMMIT$ decision.
+If not (1) and not (2), return an $ABORT$ decision if timed out; otherwise, return a $PENDING$ decision.
 
 \subsubsection{Database} \label{database}
-The database component is responsible for committing sub-transactions in each cohort. It abstracts typical ACID databases with 5 APIs in a DatabaseAdapter 
+The database component is responsible for committing sub-transactions in each cohort and abstracts typical ACID databases with 5 APIs in a DatabaseAdapter:
 \begin{itemize}
@@
 
-We require the underlying database to be ACID because the database can handle locks and transactions natively without making cohorts implementing complicated locking logic for Put and Get operations.
-But for databases that only support weak snapshot isolation (e.g.
-LMDB [4]), the cohort only needs a single write lock to prevent other write transactions to be started to make sure that transactions are serialized.
+We require the underlying database to be ACID because the database can handle locks and transactions natively without requiring cohorts to implement complicated locking logic for Put and Get operations.
+However, for databases that only support weak snapshot isolation (e.g.
+LMDB [4]), the cohort needs a single write lock to prevent other write transactions to be started to make sure that transactions are serialized.
 
 \subsubsection{Coordinator} \label{coordinator}
-The Coordinator exposes two APIs to the clients
+The coordinator exposes two APIs to the clients:
 \begin{itemize}
@@
 
-When a coordinator receives a \texttt{\\CommitAtomicTransaction} call from the client, it will first decompose the transaction into sub-transactions, which will be executed in the cohorts, based on the $namespace$ information.
-We shard the data key space into $namespaces$.
-Each cohort is responsible for a range of keys under the same $namespace$.
-After transaction decomposition, it will send \texttt{PrepareTransaction} requests with the sub-transaction info to the cohorts.
-At the same time, it will send \texttt{StartVoting} to the blockchain to trace the voting status from the cohorts.
-Different from the traditional 2PC protocol, our coordinator is designed to delegate voting and coordination work to the blockchain.
-From here on, it no longer handles any work unless \texttt{GetTransactionResult} is requested from the client.
+The data key space is split into $namespaces$, each of which corresponds to the single cohort responsible for all keys in that range.
+When the coordinator receives a \texttt{\\CommitAtomicTransaction} call from the client, it first sends \texttt{StartVoting} to the blockchain to trace the voting status from the cohorts.
+Then it decomposes the transaction into sub-transactions, which will be executed in the cohorts, based on the $namespace$ of each operation.
+After transaction decomposition, the coordinator sends \texttt{PrepareTransaction} requests with the sub-transactions to the cohorts.
+Different from the traditional 2PC protocol, we design our coordinator to delegate voting and coordination work to the blockchain.
+From here on, the coordinator no longer handles any work unless \texttt{GetTransactionResult} is requested from the client.
 
-When the client sends \texttt{GetTransactionResult} to the coordinator, it reads the transaction status from the blockchain.
-If the status is $PENDING$ or $ABORTED$, it will return the status directly.
-If the transaction is $COMMITTED$, it will further send requests to the cohorts to fetch the Get result as well and return the status and the get results altogether.
-If not all cohorts return the Get result (e.g.
-due to crashes or network latency), the coordinator will send a partial response to the client with the ones that responded along with and indicator that the response is not yet complete.
+When the client sends \texttt{GetTransactionResult} to the coordinator, the coordinator reads the transaction status from the blockchain.
+If the blockchain returns a PENDING or ABORTED status, the coordinator will return the status directly.
+If the transaction is COMMITTED, the coordinator will send requests to the cohorts to fetch the Get results and then return the status with the Get results.
+If not all cohorts return the Get results (e.g.
+due to crashes or network latency), the coordinator will send a partial response to the client with the ones that responded along with an indicator that the response is not yet complete.
+
+As described in the handling failures section, the coordinator does not need to persist any information to disk as all voting information is stored in the blockchain and all other information is also known by the client.
 
@@
 
-The Cohort exposes two APIs to the Coordinator
+The cohort exposes two APIs to the coordinator:
 \begin{itemize}
@@
 
-The $PrepareTransactionRequest$ contains a $transaction\_id$, a list of operations for the sub-transaction and an abort deadline.
-When a cohort receives a PrepareTransaction request from the coordinator, it takes these steps
+The $PrepareTransactionRequest$ contains a $transaction\_id$, a list of operations of the sub-transaction and a vote timeout time, after which the blockchain will not accept votes.
+When a cohort receives a \texttt{PrepareTransaction} request from the coordinator, it takes these steps:
 \begin{enumerate}
   \item Start a new thread in the thread pool.
-  \item Creates a \texttt{DatabaseAdapter} object.
-  \item Starts a transaction with 
+  \item Create a \texttt{DatabaseAdapter} object.
+  \item Start a transaction with 
   \texttt{DatabaseAdapter\\::Begin()}. 
-  \item Do a series of requested Put or Get operations in the sub-transaction. 
-If any operation is reported invalid by the database, the cohort will send an $ABORT$ ballot to the blockchain via \texttt{BlockchainAdapter::Vote()}.
-  \item If all operations succeed, it will store the got values in both an in-memory hashmap as well as persistently on disk keyed by the $transaction\_id$ to prepare for future \texttt{GetTransactionResult}  calls.
-  \item Set a timer based on the input abort deadline.
-  \item If the timer reaches the deadline, it will send a request \texttt{BlockchainAdapter::\\GetVotingDecision(transaction\_id)} to the blockchain. 
+  \item Acquire the relevant locks for the sub-transaction. 
+  \item Execute a series of requested Put or Get operations in the sub-transaction. 
+If any operation is reported invalid by the database or locks cannot be acquired in time, the cohort will send an $ABORT$ ballot to the blockchain via \texttt{BlockchainAdapter::Vote()}.
+  \item If all operations succeed, it will store the retrieved Get values in an in-memory hashmap as well as persistently on disk keyed by the $transaction\_id$ to prepare for future \texttt{GetTransactionResult} calls.
+  \item Set a timer based on the vote timeout time.
+  \item If the timer reaches the vote timeout time, it will send a request \texttt{BlockchainAdapter::\\GetVotingDecision(transaction\_id)} to the blockchain. 
   \begin{enumerate}
@@
 
-When the cohort receives a \texttt{GetTransaction\-Result} request, if it has committed the transaction, it simply reads the in-memory hashmap based on the $transaction\_id$ in the request and returns the got values as response.
+When the cohort receives a \texttt{GetTransaction\-Result} request and has committed the transaction, it simply reads the in-memory hashmap based on the $transaction\_id$ in the request and returns the got values as response.
 If it has not yet received a commit or abort response from the blockchain, it returns a pending response instead.
@@
 \begin{enumerate}
-\item Client composes a transaction request and passes it to Coordinator through \texttt{\\CommitAomticTransaction}.
-Coordinator returns a transaction ID for the client to look up the transaction result later.
-\item Coordinator does the followings in parallel: (a) Coordinator decomposes the transaction into sub-transaction and distributes them to the corresponding cohort through \texttt{PrepareTransaction}.
-(b) Coordinator asks Blockchain to trace votes for the new transaction through \texttt{StartVoting}.
-\item Cohort executes a sub-transaction by forwarding the operations to the database up to the point before \texttt{Commit}. Cohort writes an entry to both the undo and redo log.
-\item Cohort votes to Blockchain through \texttt{Vote} if it wants to commit or abort the transaction.
-\item Cohort sleeps until the expected transaction timeout time and gets the voting decision from Blockchain through \texttt{GetVotingDecision}.
-\item If the voting decision is to commit the transaction, cohort commits the transaction onto the database and releases acquired locks; otherwise, undoes the transaction using the undo log and releases acquired locks.
-If a transaction is committed, the cohort writes the \texttt{GET} operation results to disk.
-\item (Independent to the previous steps) Client has waited until the timeout time and requests the transaction result from Coordinator through \texttt{GetTransactionResult}, using the transaction ID given in step 1.
-\item Coordinator gets (1) the voting decision of the inquiry transaction from Blockchain through \texttt{GetTransactionResult} and (2) the results from the memory or the disk written by the cohort.
+\item The client composes a transaction request and passes it to the coordinator through \texttt{\\CommitAomticTransaction}.
+The coordinator returns a transaction ID for the client to look up the transaction result later.
+\item The coordinator asks the blockchain to trace votes for the new transaction through \texttt{StartVoting}.
+\item The coordinator decomposes the transaction into sub-transactions and distributes them to the corresponding cohort through \texttt{PrepareTransaction}.
+\item Each cohort executes a sub-transaction by acquiring locks and forwarding the operations to the database up to the point before Commit().
+The cohort writes an entry to the redo log to recover in case it crashes.
+Additionally, the cohort writes the \texttt{GET} operation results to disk for faster responses after a crash and recovery.
+\item The cohort votes to the blockchain through \texttt{Vote} indicating that it wants to commit or abort the transaction.
+\item The cohort waits until the expected transaction timeout time and gets the voting decision from the blockchain through \texttt{GetVotingDecision}.
+\item If the voting decision is to commit the transaction, cohort commits the sub-transaction onto the database and releases any acquired locks; otherwise, it aborts the sub-transaction in the database and releases any acquired locks.
+\item (Independent of the previous steps) The client waits until the timeout time and requests the transaction result from the coordinator through \texttt{GetTransactionResult}, using the transaction ID given in step 1.
+\item The coordinator gets (1) the voting decision of the inquiry transaction from the blockchain through \texttt{GetTransactionResult} and (2) the GET operation results from each cohort.
 \end{enumerate}
@@
 As described above, one of the main benefits of our system compared to traditional 2PC implementations is improved resiliency.
-By delegating the voting and coordination to the blockchain, our system will not block on any one machine that crashes or has a slow or partitioned network.
+By delegating the voting and coordination to the blockchain, our system does not block on any one machine that crashes or has a partitioned network.
 
-\subsubsection{Crash of coordinator}
+\subsubsection{Crash of the coordinator}
 The state of the coordinator includes only the list of which cohorts are involved in each active transaction.
-Thus if the coordinator crashes, when it recovers, it will respond to client requests for the Get responses of the transaction by indicating that it is unable to find the transaction.
-As a result, the client can directly ask the cohorts what the Get responses of the transaction are.
+Thus if the coordinator crashes, when it recovers, it will respond to client requests for the \texttt{GET} responses of the transaction by indicating that it is unable to find the transaction.
+As a result, the client can directly ask the cohorts what the \texttt{GET} responses of the transaction are.
 Unlike in traditional 2PC, where the coordinator stores how each cohort has voted, in our design, this information is stored in the blockchain.
@@
 
-\subsubsection{Crash of cohort}
-Our system is also resilient to crashes of cohorts, as these crashes do not cause any other cohorts to block for longer than the $vote\_timeout\_time$.
-If a cohort crashes before sending a vote to commit, then the blockchain will wait until the $vote\_timeout\_time$ and all cohorts will abort the transaction.
+\subsubsection{Crash of a cohort}
+Our system is also resilient to crashes of cohorts, as these crashes do not cause any other cohorts to block for longer than the vote timeout time.
+If a cohort crashes before sending a vote to commit, then the blockchain will wait until the vote timeout time and all cohorts will abort the transaction.
 If a cohort crashes after voting to commit and the blockchain decides that the transaction should commit, then the cohort will replay all transactions that committed while it crashed and recovered.
-To replay the transactions, the cohort will acquire the same locks it had for the transactions and send the relevant put and get requests to the database and commit the transactions in the database.
+To replay the transactions, the cohort will acquire the same locks it had for the transactions and send the relevant Put and Get requests to the database and commit the transactions in the database.
 Since it has already computed the response, if the coordinator asks for the get responses for any of these transactions, the cohort can safely return them, even if they have not yet committed in the database.
-This can be implemented in the future but we decided not to include it in our first version.
+This can be implemented in the future, but we decided not to include it in our first version.
 
@@
 According to [this article], it takes around 15 seconds to 5 minutes to process a blockchain transaction on Ethereum, which indicates that the blockchain is relatively slow compared to other components within our system.
-
-A slow blockchain affects the number of COMMIT votes to collect before the timeout time, which leads to more transactions to abort eventually; however, the blockchain slowness doesn’t affect the promised atomicity.
+A slow blockchain affects the number of commit votes to collect before the vote timeout time, which leads to more transactions to abort eventually; however, the blockchain slowness doesn’t affect the promised atomicity.
 The client or the coordinator can specify a longer timeout time if needed.
@@
 \section{Implementation}
-In this section we first describe the implementation of our overall system. Then we’ll illustrate the implementation of individual components.
+In this section we first describe the implementation of our overall system. Then, we illustrate the implementation of individual components.
 
 \subsection{Environment}
-We implement the coordinator and the cohorts with about 2800 lines of C++ code.
-For communications between components, we use gPRC and ProtoBuf.
-For providing the extensibility of swapping different database and blockchain implementations, we provide generic interfaces.
+We implemented the coordinator and the cohorts with about 2800 lines of C++ code.
+For communications between components, we used gPRC and ProtoBuf.
+For providing the extensibility of swapping different database and blockchain implementations, we provided generic interfaces.
 We tested end-to-end behavior on a single machine with different processes simulating each component.
@@
 We designed the DatabaseAdapter interface and implemented the methods with LMDB primitives.
-LMDB provides process-oriented functions like \texttt{open()}, \texttt{begin()}, \texttt{commit()}, \texttt{put()}, \texttt{get()}.
+LMDB provides process-oriented functions like \texttt{Open()}, \texttt{Begin()}, \texttt{Commit()}, \texttt{Put()}, \texttt{Get()}.
 Our object-oriented C++ implementation has almost 500 lines of code.
-We chose LMDB because it is fast, lightweight and easy to test in memory.
-But the interface is extensible to other databases like SQLite [6] as well.
+We chose LMDB because it is fast, lightweight and easy to test in memory; however, the interface is also extensible to other databases like SQLite [6].
 
@@
 One optimization we used for the cohort is that the \texttt{PrepareTransactionRequest} includes an indicator if this is the only cohort for the transaction.
-In such cases, we can avoid the expensive requests to the blockchain and the cohort can immediately commit a valid transaction as soon as it finishes preparing it in the database.
+In such cases, we avoid the expensive requests to the blockchain and the cohort can immediately commit a valid transaction as soon as it finishes preparing it in the database.
 This is safe because it knows that no other cohort can abort the transaction and that it is guaranteed to be atomic within its own database.
@@
 However, if a coordinator crashes, then clients cannot ask other coordinators for the Get responses of a transaction, since those coordinators will not know how to find the cohorts for the transaction.
-As part of a future effort, we could have the coordinators send each other the cohorts for each transaction so that they can respond to client requests even if the original coordinator for that request crashes.
+As part of a future effort, we could have the coordinators send each other the cohorts for each transaction so they can respond to client requests even if the original coordinator for that request crashes.
 

### 2022-06-03 paper/paper.tex
Revised. Saved for backup.

@@
 
-\author{\IEEEauthorblockN{Benjamin Jacob Marks}
+\author{\IEEEauthorblockN{Benjamin Marks}
   \IEEEauthorblockA{\textit{Stanford University} \\
@@
 \begin{IEEEkeywords}
-atomic commit, blockchain, smart contract, distributed system, database transaction
+  Atomic Commit, Blockchain, Smart Contract, Distributed System, Database Transaction.
 \end{IEEEkeywords}
@@
 It ensures that the transaction can be either committed or aborted at all distributed machines.
-However, blocking is proven [3] to be to be inevitable even in
-synchronous systems where bounds on delays can be reliably estimated.
+However, blocking is proven \cite{} to be to be inevitable even in synchronous systems where bounds on delays can be reliably estimated.
 
-Since the invention of Bitcoin, the underlying technology, the blockchain, not only gained attraction from  cryptocurrency systems like Ethereum but also academic research about computer systems.
-At its core, blockchain is a protocol of reaching consensus by using Proof-of-Work or Proof-of-Stake [//TODO: reference goes here].
-Previous work [1] has shown that synchronous blockchain can increase the reliability of the original 2PC protocol and also make it non-blocking when various kinds of failures happen.
+Since the invention of Bitcoin, the underlying blockchain technology has not only gained attraction from  cryptocurrency systems like Ethereum \cite{} but also academic research about computer systems.
+At its core, blockchain  is a protocol of reaching consensus by using Proof-of-Work.
+Previous work \cite{} has shown that a synchronous blockchain can increase the reliability of the original 2PC protocol and also make it non-blocking when various kinds of failures happen.
 
@@
 These contributions make 2PC with blockchain work in production and at scale.
-Compared to the traditional version of 2PC protocol [2], our proposed approach increases extensibility, scalability and reliability.
+Compared to the traditional version of 2PC protocol \cite{}, our proposed approach increases extensibility, scalability and reliability.
 
@@
 We can plug any blockchain, regardless of whether it is public/permissionless or private/permissioned, into our blockchain interface which exposes \texttt{StartVoting()}, \texttt{Vote()}, \texttt{GetVotingDecision()}.
-We can also plug any database system that has ACID support into our transaction delegation interface, which implements common transaction operations - \texttt{Begin()}, \texttt{Commit()}, \texttt{Abort()}, \texttt{Get()}, \texttt{Put()}.
+We can also plug any database system that has ACID support into our transaction delegation interface which implements common transaction operations - \texttt{Begin()}, \texttt{Commit()}, \texttt{Abort()}, \texttt{Get()}, \texttt{Put()}.
 
 \subsubsection{Scalability}
-We design our coordinator  to be light-weight to handle high throughput and horizontally scalable.
+We design our coordinator to be light-weight to handle high throughput and to be horizontally scalable.
 This is achieved by delegating coordination and voting work to the blockchain.
@@
 Our system has better reliability than the traditional 2PC protocol for two reasons.
-First, the system leverages blockchain to maintain the coordination state.
+First, the system leverages the blockchain to maintain the coordination state.
 The stateless coordinator can easily recover as long as it knows the participating cohorts.
 Second, The coordinator is stateless and horizontally scalable.
-We can keep multiple jobs running concurrently and place a load-balancing layer on top of it.
+We can keep multiple jobs running concurrently and place a load-balancing layer on top of it to be tolerant of individual coordinator node failures.
 
@@
 \subsection{Architecture and Components} \label{components}
-We define a transaction as a list of operations starting with \texttt{BEGIN} followed by multiple \texttt{GET} or \texttt{PUT} and ending with \texttt{COMMIT}.
+
+We define a transaction as a list of operations starting with \texttt{Begin} followed by multiple \texttt{Get} or \texttt{Put} and ending with \texttt{Commit}.
 The client creates a transaction and requests the coordinator to commit.
 The coordinator distributes the transaction to the corresponding cohorts to prepare and then offloads the remaining 2PC responsibility to the blockchain.
-Each cohort prepares the transaction locally, votes to the blockchain if ready, checks on Blockchain to see if a transaction should be committed, then finally commits the transaction to its database.
+Each cohort prepares the transaction locally, votes to the blockchain if ready, checks on the blockchain to see if a transaction should be committed, and finally commits the transaction to its database.
 
-The architecture is similar to the traditional 2PC architecture where we have one coordinator and multiple cohorts - except each cohort talks to the blockchain directly to vote and to look up the voting decision as the coordinator offload the 2PC voting responsibility to the blockchain.
+The architecture is similar to the traditional 2PC architecture where we have one coordinator and multiple cohorts - except each cohort talks to the blockchain directly to vote and to look up the voting decision as the coordinator offloads the 2PC voting responsibility to the blockchain.
 
-We describe details of each component and each in the remaining \ref{components}, the complete flow of handling a successful transaction in \ref{txnflow}, and how we handle failures in \ref{failurecases}.
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.4\textwidth]{./diagram.png}
-\caption{Fig 1. This diagram displays the main components within the whole architecture and example steps of successfully committing a transaction.}
-\label{fig:Panel}
-\end{figure}
+We describe the details of each component in the remaining \ref{components}, the complete flow of handling a successful transaction in \ref{tnxflow}, and how we handle failures in \ref{failurecases}.
+
+% \begin{figure}[h]
+  % \centering
+  % \includegraphics[width=0.4\textwidth]{./diagram.png}
+% \caption{Fig 1. This diagram displays the main components within our architecture and example steps of successfully committing a transaction.}
+  % \label{fig:Panel}
+% \end{figure}
 
 \subsubsection{Blockchain} \label{blockchain}
+
+The blockchain component performs a 2PC protocol, so it’s responsible for tracking voting states of each transaction and making decisions of whether to commit or abort.
+The implementation uses Ethereum smart contracts to store the followings states of each transaction (indexed by the transaction ID) internally:
+
 The blockchain component performs a 2PC protocol, so it’s responsible for tracking voting states of each transaction and making decisions of committing.
 The implementation uses Ethereum smart contracts to store the followings states of each transaction (indexed by the transaction ID) internally:
-$transaction\_state$ stores voting states of each cohort of a transaction.
-$transaction\_config$ stores the transaction configure of each transaction including the number of cohorts and the vote timeout time.
 
-We designed a \texttt{\\BlockchainAdapter} interface, which exposes the following APIs externally to the coordinator and the cohort:
-\texttt{StartVoting(\\transaction\_id, cohorts, vote\_timeout\_time)} Called by the coordinator to start voting on a transaction.
-$cohorts$ is the number of cohorts that will vote on this transaction later.
-$vote\_timeout\_time$ is the absolute time on the blockchain when the 2PC stops taking votes.
-\texttt{Vote(transaction\_id, cohort\_id, ballot)} Called by the cohort to vote $COMMIT$ or $ABORT$ on a transaction before timeout.
-\texttt{GetVotingDecision(transaction\_id)} Called by any component to check the voting state on a transaction.
+\begin{itemize}
+  \item \texttt{transaction\_state} stores voting states of each cohort of a transaction.
+  \item \texttt{transaction\_config} stores the transaction configure of each transaction including the number of cohorts and the vote timeout time.
+\end{itemize}
 
-\texttt{StartVoting} initializes the states for a transaction and \texttt{Vote} updates the voting status of a transaction. \texttt{GetVotingDecision} makes the decision by following the below logic:
-If any $ABORT$ vote is received, return an $ABORT$ decision.
-If all $COMMIT$ votes are received, return a $COMMIT$ decision.
-If not (1) and not (2), return an $ABORT$ decision if timed out; otherwise, return a $PENDING$ decision.
+We designed a \texttt{BlockchainAdapter} interface, which exposes the following APIs externally to the coordinator and the cohort:
+\begin{itemize}
+  \item \texttt{StartVoting(transaction\_id, cohorts, vote\_timeout\_time)} Called by the coordinator to start voting on a transaction.
+\texttt{cohorts} is the number of cohorts that will vote on this transaction later.
+\texttt{vote\_timeout\_time} is the absolute time on the blockchain when the 2PC stops taking votes.
+  \item \texttt{Vote(transaction\_id, cohort\_id, ballot)} Called by the cohort to vote \texttt{COMMIT} or \texttt{ABORT} on a transaction before timeout.
+  \item \texttt{GetVotingDecision(transaction\_id)} Called by any component to check the voting state of a transaction.
+\end{itemize}
+
+\texttt{StartVoting()} initializes the states for a transaction and \texttt{Vote()} updates the voting status of a transaction.
+\texttt{GetVotingDecision()} makes the decision by following the below logic:
+\begin{enumerate}
+  \item If any \texttt{ABORT} vote is received, return an \texttt{ABORT} decision.
+  \item If all \texttt{COMMIT} votes are received, return a \texttt{COMMIT} decision.
+  \item If not (1) and not (2), return an \texttt{ABORT} decision if the blockchain time has passed the vote timeout time; otherwise, return a \texttt{PENDING} decision.
+\end{enumerate}
 
 \subsubsection{Database} \label{database}
-The database component is responsible for committing sub-transactions in each cohort and abstracts typical ACID databases with 5 APIs in a DatabaseAdapter:
+The database component is responsible for committing sub-transactions in each cohort and abstracts typical ACID databases with 5 APIs in a \texttt{DatabaseAdapter:} 
 \begin{itemize}
-  \item \texttt{Begin()}:  starts a transaction
-  \item \texttt{Commit()}: commits a started transaction
-  \item \texttt{Abort()}: aborts a started transaction
-  \item \texttt{Put(key, value)}: puts value in a row whose primary key is key 
-  \item \texttt{Get(key)}: gets the value from the row whose primary key is key 
+  \item \texttt{Begin()} starts a transaction.
+  \item \texttt{Commit()} commits a started transaction.
+  \item \texttt{Abort()} aborts a started transaction.
+  \item \texttt{Put(key, value)} puts value in a row whose primary key is key.
+  \item \texttt{Get(key)} gets the value from the row whose primary key is key.
 \end{itemize}
 
-We require the underlying database to be ACID because the database can handle locks and transactions natively without requiring cohorts to implement complicated locking logic for Put and Get operations.
+We require the underlying database to be ACID because the database can handle locks and transactions natively without requiring cohorts to implement complicated locking logic for \texttt{Put} and \texttt{Get} operations.
 However, for databases that only support weak snapshot isolation (e.g.
-LMDB [4]), the cohort needs a single write lock to prevent other write transactions to be started to make sure that transactions are serialized.
+LMDB \cite{}), the cohort needs a single write lock to prevent other write transactions to be started to make sure that transactions are serialized.
 
@@
 \begin{itemize}
-  \item \texttt{CommitAtomicTransaction(
-    CommitAtomicTransactionRequest, CommitAtomicTransactionResponse)}
-  \item \texttt{GetTransactionResult(
-    GetTransactionResultRequest, GetTransactionResultResponse)}
+  \item \texttt{CommitAtomicTransaction(CommitAtomicTransactionRequest, CommitAtomicTransactionResponse)}
+  \item \texttt{GetTransactionResult(GetTransactionResultRequest, GetTransactionResultResponse)}
 \end{itemize}
@@
 
+The data key space is split into \texttt{namespaces}, each of which corresponds to the single cohort responsible for all keys in that range.
+When the coordinator receives a \texttt{CommitAtomicTransaction()} call from the client, it first sends \texttt{StartVoting()} to the blockchain to track the voting status from the cohorts.
+Then it decomposes the transaction into sub-transactions, which will be executed in the cohorts, based on the namespace of each operation.
+After transaction decomposition, the coordinator sends \texttt{PrepareTransaction()} requests with the sub-transactions to the cohorts.
+Different from the traditional 2PC protocol, we design our coordinator to delegate voting and coordination work to the blockchain.
+From here on, the coordinator no longer handles any work unless \texttt{GetTransactionResult()} is requested from the client.
+
+When the client sends \texttt{GetTransactionResult()} to the coordinator, the coordinator reads the transaction status from the blockchain.
+If the blockchain returns a \texttt{PENDING} or \texttt{ABORTED} status, the coordinator will return the status directly.
+If the transaction is \texttt{COMMITTED}, the coordinator will send requests to the cohorts to fetch the \texttt{Get} results and then return the status with the \texttt{Get} results.
+If not all cohorts return the \texttt{Get} results (e.g.
+due to crashes or network latency), the coordinator will send a partial response to the client with the ones that responded along with an indicator that the response is not yet complete.
+
+As described in \ref{}, the coordinator does not need to persist any information to disk as all voting information is stored in the blockchain and all other information (e.g.
+list of cohorts involved in the transaction) is also known by the client.
+
 \subsubsection{Cohort} \label{cohort}
-
 The cohort exposes two APIs to the coordinator:

### 2022-06-04 paper/paper.tex
Update draft (citation hasn't completed yet).

@@
 \usepackage{xcolor}
+\usepackage[htt]{hyphenat}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
 T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
 \begin{document}
@@
 
-We describe the details of each component in the remaining \ref{components}, the complete flow of handling a successful transaction in \ref{tnxflow}, and how we handle failures in \ref{failurecases}.
+We describe the details of each component in the remaining \ref{components}, the complete flow of handling a successful transaction in \ref{txnflow}, and how we handle failures in \ref{failurecases}.
 
-% \begin{figure}[h]
-  % \centering
-  % \includegraphics[width=0.4\textwidth]{./diagram.png}
-% \caption{Fig 1. This diagram displays the main components within our architecture and example steps of successfully committing a transaction.}
-  % \label{fig:Panel}
-% \end{figure}
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{./diagram.pdf}
+  \caption{This diagram displays the main components within our architecture and example steps of successfully committing a transaction.}
+  \label{fig:Panel}
+\end{figure*}
 
@@
 \begin{itemize}
-  \item \texttt{CommitAtomicTransaction(CommitAtomicTransactionRequest, CommitAtomicTransactionResponse)}
-  \item \texttt{GetTransactionResult(GetTransactionResultRequest, GetTransactionResultResponse)}
+  \item \texttt{CommitAtomicTransaction( CommitAtomicTransactionRequest, CommitAtomicTransactionResponse)}
+  \item \texttt{GetTransactionResult( GetTransactionResultRequest, GetTransactionResultResponse)}
 \end{itemize}
@@
 \begin{itemize}
-  \item \texttt{PrepareTransaction(
-    PrepareTransactionRequest, PrepareTransactionResponse)}
-  \item \texttt{GetTransactionResult(
-    GetTransactionResultRequest, GetTransactionResultResponse)}
+  \item \texttt{PrepareTransaction(\\PrepareTransactionRequest, PrepareTransactionResponse)}
+  \item \texttt{GetTransactionResult(\\GetTransactionResultRequest, GetTransactionResultResponse)}
 \end{itemize}
 
-The $PrepareTransactionRequest$ contains a $transaction\_id$, a list of operations of the sub-transaction and a vote timeout time, after which the blockchain will not accept votes.
-When a cohort receives a \texttt{PrepareTransaction} request from the coordinator, it takes these steps:
+The \texttt{PrepareTransactionRequest()} contains a \texttt{transaction\_id}, a list of operations of the sub-transaction and the vote timeout time, after which the blockchain will not accept votes.
+When a cohort receives a \texttt{PrepareTransaction()} request from the coordinator, it takes these steps:
 \begin{enumerate}
@@
   \item Create a \texttt{DatabaseAdapter} object.
-  \item Start a transaction with 
-    \texttt{DatabaseAdapter\\::Begin()}. 
+  \item Start a transaction with \texttt{DatabaseAdapter:: Begin()}.
   \item Acquire the relevant locks for the sub-transaction.
-  \item Execute a series of requested Put or Get operations in the sub-transaction. 
-    If any operation is reported invalid by the database or locks cannot be acquired in time, the cohort will send an $ABORT$ ballot to the blockchain via \texttt{BlockchainAdapter::Vote()}.
-  \item If all operations succeed, it will store the retrieved Get values in an in-memory hashmap as well as persistently on disk keyed by the $transaction\_id$ to prepare for future \texttt{GetTransactionResult} calls.
-  \item Set a timer based on the vote timeout time.
+  \item Execute a series of requested \texttt{Put} or \texttt{Get} operations in the sub-transaction. 
+  \item If any operation is reported invalid by the database or locks cannot be acquired in time, the cohort will send an \texttt{ABORT} ballot to the blockchain via \texttt{BlockchainAdapter::Vote()}.
+  \item If all operations succeed, it will store the retrieved Get values in an in-memory hashmap as well as persistently on disk keyed by the \texttt{transaction\_id} to prepare for future \texttt{GetTransactionResult()} calls.
+  \item Set a timer based on the input vote timeout time.
   \item If the timer reaches the vote timeout time, it will send a request \texttt{BlockchainAdapter::\\GetVotingDecision(transaction\_id)} to the blockchain. 
     \begin{enumerate}
-      \item If the blockchain states that the transaction should be committed, it will call \texttt{DatabaseAdapter::Commit()}.
-      \item If the blockchain states that the transaction should be aborted, it will call \texttt{DatabaseAdapter::Abort()}.
+      \item If the blockchain states that the transaction should be committed, it will call \texttt{DatabaseAdapter::Commit()} and release any acquired locks.
+  \item If the blockchain states that the transaction should be aborted, it will call \texttt{DatabaseAdapter::Abort()} and release any acquired locks.
+  \end{enumerate}
 \end{enumerate}
 
-\end{enumerate}
-
-When the cohort receives a \texttt{GetTransaction\-Result} request and has committed the transaction, it simply reads the in-memory hashmap based on the $transaction\_id$ in the request and returns the got values as response.
+When the cohort receives a \texttt{GetTransactionResult()} request and has committed the transaction, it simply reads the in-memory hashmap based on the \texttt{transaction\_id} in the request and returns the got values as response.
 If it has not yet received a commit or abort response from the blockchain, it returns a pending response instead.
@@
 \subsection{System Interactions (a.k.a Life of a Transaction)} \label{txnflow}
+
 This section describes the end-to-end life of a transaction without handling failures.
@@
 \begin{enumerate}
-  \item The client composes a transaction request and passes it to the coordinator through \texttt{\\CommitAomticTransaction}.
+  \item The client composes a transaction request and passes it to the coordinator through \texttt{CommitAomticTransaction()}.
 The coordinator returns a transaction ID for the client to look up the transaction result later.
-  \item The coordinator asks the blockchain to trace votes for the new transaction through \texttt{StartVoting}.
-  \item The coordinator decomposes the transaction into sub-transactions and distributes them to the corresponding cohort through \texttt{PrepareTransaction}.
+  \item The coordinator asks the blockchain to trace votes for the new transaction through \texttt{StartVoting()}.
+  \item The coordinator decomposes the transaction into sub-transactions and distributes them to the corresponding cohort through \texttt{PrepareTransaction()}.
   \item Each cohort executes a sub-transaction by acquiring locks and forwarding the operations to the database up to the point before Commit().
@@
 Additionally, the cohort writes the \texttt{GET} operation results to disk for faster responses after a crash and recovery.
-  \item The cohort votes to the blockchain through \texttt{Vote} indicating that it wants to commit or abort the transaction.
-  \item The cohort waits until the expected transaction timeout time and gets the voting decision from the blockchain through \texttt{GetVotingDecision}.
-  \item If the voting decision is to commit the transaction, cohort commits the sub-transaction onto the database and releases any acquired locks; otherwise, it aborts the sub-transaction in the database and releases any acquired locks.
-  \item (Independent of the previous steps) The client waits until the timeout time and requests the transaction result from the coordinator through \texttt{GetTransactionResult}, using the transaction ID given in step 1.
-  \item The coordinator gets (1) the voting decision of the inquiry transaction from the blockchain through \texttt{GetTransactionResult} and (2) the GET operation results from each cohort.
+  \item The cohort votes to the blockchain through \texttt{Vote()} indicating that it wants to commit or abort the transaction.
+  \item The cohort waits until the expected transaction vote timeout time and gets the voting decision from the blockchain through \texttt{GetVotingDecision()}.
+  \item If the voting decision is to commit the transaction, the cohort commits the sub-transaction onto the database and releases any acquired locks; otherwise, it aborts the sub-transaction in the database and releases any acquired locks.
+  \item (Independent of the previous steps) The client waits until the vote timeout time and requests the transaction result from the coordinator through \texttt{GetTransactionResult()}, using the transaction ID given in step 1.
+  \item The coordinator gets (1) the voting decision of the inquiry transaction from the blockchain through \texttt{GetTransactionResult()} and (2) the \texttt{GET} operation results from each cohort.
 \end{enumerate}
@@
 The state of the coordinator includes only the list of which cohorts are involved in each active transaction.
-Thus if the coordinator crashes, when it recovers, it will respond to client requests for the \texttt{GET} responses of the transaction by indicating that it is unable to find the transaction.
-As a result, the client can directly ask the cohorts what the \texttt{GET} responses of the transaction are.
+Thus if the coordinator crashes, when it recovers, it will respond to client requests for the \texttt{Get} responses of the transaction by indicating that it is unable to find the transaction.
+As a result, the client can directly ask the cohorts what the \texttt{Get} responses of the transaction are.
 Unlike in traditional 2PC, where the coordinator stores how each cohort has voted, in our design, this information is stored in the blockchain.
-Our system is resilient to such crashes, as the transaction does not need to block on the coordinator’s recovery.
+Our system is resilient to such crashes, as the transaction does not need to block on the coordinator’s recovery, and the cohorts can commit or abort even while the coordinator is unresponsive.
 
@@
 If a cohort crashes after voting to commit and the blockchain decides that the transaction should commit, then the cohort will replay all transactions that committed while it crashed and recovered.
-To replay the transactions, the cohort will acquire the same locks it had for the transactions and send the relevant Put and Get requests to the database and commit the transactions in the database.
-Since it has already computed the response, if the coordinator asks for the get responses for any of these transactions, the cohort can safely return them, even if they have not yet committed in the database.
+To replay the transactions, the cohort will acquire the same locks it had for the transactions and send the relevant \texttt{Put} and \texttt{Get} requests to the database and commit the transactions in the database.
+Since it has already computed the response, if the coordinator asks for the \texttt{Get} responses for any of these transactions, the cohort can safely return them, even if they have not yet committed in the database.
 This can be implemented in the future, but we decided not to include it in our first version.
@@
 We use Ethereum for our blockchain implementation.
-According to [this article], it takes around 15 seconds to 5 minutes to process a blockchain transaction on Ethereum, which indicates that the blockchain is relatively slow compared to other components within our system.
-A slow blockchain affects the number of commit votes to collect before the vote timeout time, which leads to more transactions to abort eventually; however, the blockchain slowness doesn’t affect the promised atomicity.
-The client or the coordinator can specify a longer timeout time if needed.
+According to \cite{}, it takes around 15 seconds to 5 minutes to process a blockchain transaction on Ethereum, which indicates that the blockchain is relatively slow compared to other components within our system.
+A slow blockchain increases the likelihood that not all commit votes can be collected before the vote timeout time, which leads to more transactions to abort eventually; however, the blockchain slowness doesn’t affect the promised atomicity.
+The client or the coordinator can specify a longer vote timeout time if needed.
 
 \section{Implementation}
-In this section we first describe the implementation of our overall system. Then, we illustrate the implementation of individual components.
+In this section we first describe the implementation of our overall system.
+Then, we illustrate the implementation of individual components.
+The source code of our implementation is available at \cite{} licensed under LGPL.
+Our implementation depends on these libraries: \cite{}, ...
 
@@
 We implemented the coordinator and the cohorts with about 2800 lines of C++ code.
-For communications between components, we used gPRC and ProtoBuf.
+For communications between components, we used gRPC \cite{} and ProtoBuf \cite{}.
 For providing the extensibility of swapping different database and blockchain implementations, we provided generic interfaces.
-We tested end-to-end behavior on a single machine with different processes simulating each component.
+We tested end-to-end behaviors on a single machine with different processes simulating each component.
 
 \subsubsection{Blockchain}
-We wrote our two-phase-commit smart contract with Solidity and developed it with Truffle [\url{https://trufflesuite.com/}].
-Truffle provides the development environment for us to compile smart contracts, test it, and deploy it locally for debugging.
-Once we’ve deployed our smart contract, the coordinator and the cohorts can interact with the contract through a smart contract client library.
-However, we didn’t find a proper C++ smart contract client library to use, so we use a JavaScript library (web.js[\url{https://web3js.readthedocs.io/en/v1.7.3/}]) instead and add an additional gRPC layer to translate to C++.
+We wrote our two-phase-commit smart contract with Solidity and developed it with Truffle \cite{}.
+Truffle provided the development environment for us to compile smart contracts, test it, and deploy it locally for debugging.
+Once we deployed our smart contract, the coordinator and the cohorts interacted with the contract through a smart contract client library.
+However, we didn’t find a proper C++ smart contract client library to use, so we used a JavaScript library (web.js[ref]) instead and added an additional gRPC layer to translate to C++.
 
-While testing in the Truffle development, we learned the gas used for successfully committing a transaction with two cohorts is around 0.33 USD [\href{https://automatedwebtools.com/gwei-to-usd-calculator/}{reference}] (see \ref{appendix} for breakdowns).
+While testing in the Truffle development, we learned the gas used for successfully committing a transaction with two cohorts is around 0.33 USD \cite{} (see \ref{appendix} for breakdowns).
 
 \subsubsection{Database}
-We designed the DatabaseAdapter interface and implemented the methods with LMDB primitives.
-LMDB provides process-oriented functions like \texttt{Open()}, \texttt{Begin()}, \texttt{Commit()}, \texttt{Put()}, \texttt{Get()}.
+We designed the \texttt{DatabaseAdapter} interface and implemented the methods with LMDB primitives.
+LMDB provides process-oriented functions like \texttt{open()}, \texttt{begin()}, \texttt{commit()}, \texttt{put()}, \texttt{get()}.
 Our object-oriented C++ implementation has almost 500 lines of code.
-We chose LMDB because it is fast, lightweight and easy to test in memory; however, the interface is also extensible to other databases like SQLite [6].
+We chose LMDB because it is fast, lightweight and easy to test in memory.
+The interface is extensible to other databases like SQLite \cite{}.
 
@@
 \subsubsection{Cohort}
-One optimization we used for the cohort is that the \texttt{PrepareTransactionRequest} includes an indicator if this is the only cohort for the transaction.
+One optimization we used for the cohort is that the \texttt{PrepareTransactionRequest()} includes an indicator if this is the only cohort for the transaction.
 In such cases, we avoid the expensive requests to the blockchain and the cohort can immediately commit a valid transaction as soon as it finishes preparing it in the database.
@@
 Our approach enables the client to tradeoff between latency and the likelihood of committing a transaction.
-When the client sets a longer time for timeout, it’s more likely to collect sufficient votes from all cohorts to make a COMMIT decision but may have higher latency if a cohort crashes.
+When the client sets a longer time for the vote timeout time, it’s more likely to collect sufficient votes from all cohorts to make a commit decision but may have higher latency if a cohort crashes.
 There is a limit to the minimum timeout, which can be no shorter than the time it takes to finalize a new blockchain block.
 
-As part of a future effort, we can extend the system to tradeoff between further latency reductions and safety by having cohorts send their votes to the coordinator as well as to the blockchain.
+As part of a future effort, we can extend the system to tradeoff between further latency reductions and safety by having cohorts send their votes to the coordinator and to the blockchain.
 If the coordinator receives commit votes from all cohorts, it can update the decision in the blockchain and tell all cohorts to commit.

### 2022-06-04 paper/paper.tex
Fix few duplicated text.

@@
 It ensures that the transaction can be either committed or aborted at all distributed machines.
-However, blocking is proven \cite{} to be to be inevitable even in synchronous systems where bounds on delays can be reliably estimated.
+However, blocking is proven \cite{ezhilchelvan2020non} to be to be inevitable even in synchronous systems where bounds on delays can be reliably estimated.
 
@@
 
-The blockchain component performs a 2PC protocol, so it’s responsible for tracking voting states of each transaction and making decisions of committing.
-The implementation uses Ethereum smart contracts to store the followings states of each transaction (indexed by the transaction ID) internally:
-
 \begin{itemize}
@@
 \begin{itemize}
-  \item \texttt{StartVoting(transaction\_id, cohorts, vote\_timeout\_time)} Called by the coordinator to start voting on a transaction.
+  \item \texttt{StartVoting(transaction\_id, cohorts, vote\_timeout\_time)}\\Called by the coordinator to start voting on a transaction.
 \texttt{cohorts} is the number of cohorts that will vote on this transaction later.
 \texttt{vote\_timeout\_time} is the absolute time on the blockchain when the 2PC stops taking votes.
-  \item \texttt{Vote(transaction\_id, cohort\_id, ballot)} Called by the cohort to vote \texttt{COMMIT} or \texttt{ABORT} on a transaction before timeout.
-  \item \texttt{GetVotingDecision(transaction\_id)} Called by any component to check the voting state of a transaction.
+  \item \texttt{Vote(transaction\_id, cohort\_id, ballot)}\\Called by the cohort to vote \texttt{COMMIT} or \texttt{ABORT} on a transaction before timeout.
+  \item \texttt{GetVotingDecision(transaction\_id)}\\Called by any component to check the voting state of a transaction.
 \end{itemize}
@@
 
-The data key space is split into $namespaces$, each of which corresponds to the single cohort responsible for all keys in that range.
-When the coordinator receives a \texttt{\\CommitAtomicTransaction} call from the client, it first sends \texttt{StartVoting} to the blockchain to trace the voting status from the cohorts.
-Then it decomposes the transaction into sub-transactions, which will be executed in the cohorts, based on the $namespace$ of each operation.
-After transaction decomposition, the coordinator sends \texttt{PrepareTransaction} requests with the sub-transactions to the cohorts.
-Different from the traditional 2PC protocol, we design our coordinator to delegate voting and coordination work to the blockchain.
-From here on, the coordinator no longer handles any work unless \texttt{GetTransactionResult} is requested from the client.
-
-When the client sends \texttt{GetTransactionResult} to the coordinator, the coordinator reads the transaction status from the blockchain.
-If the blockchain returns a PENDING or ABORTED status, the coordinator will return the status directly.
-If the transaction is COMMITTED, the coordinator will send requests to the cohorts to fetch the Get results and then return the status with the Get results.
-If not all cohorts return the Get results (e.g.
-due to crashes or network latency), the coordinator will send a partial response to the client with the ones that responded along with an indicator that the response is not yet complete.
-
-As described in the handling failures section, the coordinator does not need to persist any information to disk as all voting information is stored in the blockchain and all other information is also known by the client.
+As described in \ref{}, the coordinator does not need to persist any information to disk as all voting information is stored in the blockchain and all other information (e.g.
+list of cohorts involved in the transaction) is also known by the client.
 
@@
 
-As described in \ref{}, the coordinator does not need to persist any information to disk as all voting information is stored in the blockchain and all other information (e.g.
+As described in \ref{blockchain}, the coordinator does not need to persist any information to disk as all voting information is stored in the blockchain and all other information (e.g.
 list of cohorts involved in the transaction) is also known by the client.

### 2022-06-04 paper/generate.sh
Link reference and remove other duplicated line.

@@ -1,2 +1,5 @@
 #!/bin/bash
-pdflatex paper.tex
+pdflatex paper
+bibtex paper
+pdflatex paper
+pdflatex paper

### 2022-06-04 paper/paper.tex
Link reference and remove other duplicated line.
Update appendix and references titles.
Add consistent Get and Put syntax.

@@
 
-As described in \ref{}, the coordinator does not need to persist any information to disk as all voting information is stored in the blockchain and all other information (e.g.
-list of cohorts involved in the transaction) is also known by the client.
-
 The data key space is split into \texttt{namespaces}, each of which corresponds to the single cohort responsible for all keys in that range.
@@
   \item If any operation is reported invalid by the database or locks cannot be acquired in time, the cohort will send an \texttt{ABORT} ballot to the blockchain via \texttt{BlockchainAdapter::Vote()}.
-  \item If all operations succeed, it will store the retrieved Get values in an in-memory hashmap as well as persistently on disk keyed by the \texttt{transaction\_id} to prepare for future \texttt{GetTransactionResult()} calls.
+  \item If all operations succeed, it will store the retrieved \texttt{Get} values in an in-memory hashmap as well as persistently on disk keyed by the \texttt{transaction\_id} to prepare for future \texttt{GetTransactionResult()} calls.
   \item Set a timer based on the input vote timeout time.
@@
 The cohort writes an entry to the redo log to recover in case it crashes.
-Additionally, the cohort writes the \texttt{GET} operation results to disk for faster responses after a crash and recovery.
+Additionally, the cohort writes the \texttt{Get} operation results to disk for faster responses after a crash and recovery.
   \item The cohort votes to the blockchain through \texttt{Vote()} indicating that it wants to commit or abort the transaction.
@@
   \item (Independent of the previous steps) The client waits until the vote timeout time and requests the transaction result from the coordinator through \texttt{GetTransactionResult()}, using the transaction ID given in step 1.
-  \item The coordinator gets (1) the voting decision of the inquiry transaction from the blockchain through \texttt{GetTransactionResult()} and (2) the \texttt{GET} operation results from each cohort.
+  \item The coordinator gets (1) the voting decision of the inquiry transaction from the blockchain through \texttt{GetTransactionResult()} and (2) the \texttt{Get} operation results from each cohort.
 \end{enumerate}
@@
 Currently the coordinator is stateless, so it is easy to add new coordinator servers, as long as each one handles a disjoint set of transactions.
-However, if a coordinator crashes, then clients cannot ask other coordinators for the Get responses of a transaction, since those coordinators will not know how to find the cohorts for the transaction.
+However, if a coordinator crashes, then clients cannot ask other coordinators for the \texttt{Get} responses of a transaction, since those coordinators will not know how to find the cohorts for the transaction.
 As part of a future effort, we could have the coordinators send each other the cohorts for each transaction so they can respond to client requests even if the original coordinator for that request crashes.
 
-\section{Reference}
+\bibliographystyle{plain}
+\bibliography{references}
 
-\section{Appendix} \label{appendix}
+\appendix
+\label{appendix}
 \begin{table}[hbtp]

### 2022-06-04 paper/paper.tex
Update some ref.

@@
 \IEEEoverridecommandlockouts
-\usepackage{cite}
+\usepackage[sort,nocompress]{cite}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 It ensures that the transaction can be either committed or aborted at all distributed machines.
-However, blocking is proven \cite{ezhilchelvan2020non} to be to be inevitable even in synchronous systems where bounds on delays can be reliably estimated.
+However, blocking is proven \cite{skeen1981nonblocking} to be to be inevitable even in synchronous systems where bounds on delays can be reliably estimated.
 
-Since the invention of Bitcoin, the underlying blockchain technology has not only gained attraction from  cryptocurrency systems like Ethereum \cite{} but also academic research about computer systems.
+Since the invention of Bitcoin, the underlying blockchain technology has not only gained attraction from  cryptocurrency systems like Ethereum \cite{wood2014ethereum} but also academic research about computer systems.
 At its core, blockchain  is a protocol of reaching consensus by using Proof-of-Work.
-Previous work \cite{} has shown that a synchronous blockchain can increase the reliability of the original 2PC protocol and also make it non-blocking when various kinds of failures happen.
+Previous work \cite{ezhilchelvan2020non} has shown that a synchronous blockchain can increase the reliability of the original 2PC protocol and also make it non-blocking when various kinds of failures happen.
 
@@
 These contributions make 2PC with blockchain work in production and at scale.
-Compared to the traditional version of 2PC protocol \cite{}, our proposed approach increases extensibility, scalability and reliability.
+Compared to the traditional version of 2PC protocol \cite{lampson1993new}, our proposed approach increases extensibility, scalability and reliability.
 
@@
 However, for databases that only support weak snapshot isolation (e.g.
-LMDB \cite{}), the cohort needs a single write lock to prevent other write transactions to be started to make sure that transactions are serialized.
+LMDB \cite{chu2020lightning}), the cohort needs a single write lock to prevent other write transactions to be started to make sure that transactions are serialized.
 
@@
 We use Ethereum for our blockchain implementation.
-According to \cite{}, it takes around 15 seconds to 5 minutes to process a blockchain transaction on Ethereum, which indicates that the blockchain is relatively slow compared to other components within our system.
+According to \cite{stationlong}, it takes around 15 seconds to 5 minutes to process a blockchain transaction on Ethereum, which indicates that the blockchain is relatively slow compared to other components within our system.
 A slow blockchain increases the likelihood that not all commit votes can be collected before the vote timeout time, which leads to more transactions to abort eventually; however, the blockchain slowness doesn’t affect the promised atomicity.
@@
 Then, we illustrate the implementation of individual components.
-The source code of our implementation is available at \cite{} licensed under LGPL.
-Our implementation depends on these libraries: \cite{}, ...
+The source code of our implementation is available at licensed under GNU Lesser General Public License.
+Our implementation depends on the following libraries: ganache, truffle, sqlite3, grpc/grpc-js, grpc/proto-loader, web3, web3-eth-contract, absl, rule\_nodejs, gflags, glog, openssl, threadpool \cite{Shoshany2021_ThreadPool}, lmdb, and lmdbxx.
 
@@
 We chose LMDB because it is fast, lightweight and easy to test in memory.
-The interface is extensible to other databases like SQLite \cite{}.
+The interface is extensible to other databases like SQLite \cite{bhosale2015sqlite}.
 

### 2022-06-04 paper/paper.tex
Updates paper.

@@
 However, for databases that only support weak snapshot isolation (e.g.
-LMDB \cite{chu2020lightning}), the cohort needs a single write lock to prevent other write transactions to be started to make sure that transactions are serialized.
+LMDB \cite{henry2019howard}), the cohort needs a single write lock to prevent other write transactions to be started to make sure that transactions are serialized.
 
@@
 
-While testing in the Truffle development, we learned the gas used for successfully committing a transaction with two cohorts is around 0.33 USD \cite{} (see \ref{appendix} for breakdowns).
+While testing in the Truffle development, we learned the gas used for successfully committing a transaction with two cohorts is around 0.33 USD \cite{gwei} (see \ref{appendix} for breakdowns).
 

### 2022-06-04 paper/generate.sh

@@ -1,5 +1,2 @@
 #!/bin/bash
-pdflatex paper
-bibtex paper
-pdflatex paper
-pdflatex paper
+pdflatex paper && bibtex paper && pdflatex paper && pdflatex paper

### 2022-06-04 paper/paper.tex
Add citation and update diagram.

@@
 \usepackage{textcomp}
+\usepackage{url}
 \usepackage{xcolor}
@@
 The source code of our implementation is available at licensed under GNU Lesser General Public License.
-Our implementation depends on the following libraries: ganache, truffle, sqlite3, grpc/grpc-js, grpc/proto-loader, web3, web3-eth-contract, absl, rule\_nodejs, gflags, glog, openssl, threadpool \cite{Shoshany2021_ThreadPool}, lmdb, and lmdbxx.
+Our implementation depends on the following libraries: truffle \cite{github_truffle}, gganacheanache \cite{github_ganache}, sqlite3 \cite{github_sqlite3}, grpc \cite{github_grpc}, web3 \cite{github_web3}, abseil-cpp \cite{github_absl}, rule\_nodejs \cite{github_rulejs}, gflags \cite{github_gflags}, glog \cite{github_glog}, openssl \cite{github_openssl}, threadpool \cite{Shoshany2021_ThreadPool}, lmdb \cite{github_lmdb}, and lmdbxx \cite{github_lmdbxx}.
 
@@
 We implemented the coordinator and the cohorts with about 2800 lines of C++ code.
-For communications between components, we used gRPC \cite{} and ProtoBuf \cite{}.
+For communications between components, we used gRPC \cite{github_grpc} and ProtoBuf \cite{github_proto}.
 For providing the extensibility of swapping different database and blockchain implementations, we provided generic interfaces.
@@
 \subsubsection{Blockchain}
-We wrote our two-phase-commit smart contract with Solidity and developed it with Truffle \cite{}.
+We wrote our two-phase-commit smart contract with Solidity and developed it with Truffle \cite{github_truffle}.
 Truffle provided the development environment for us to compile smart contracts, test it, and deploy it locally for debugging.

### 2022-06-04 paper/generate.sh

@@ -1,2 +1,3 @@
 #!/bin/bash
+rm -f paper.aux paper.log paper.bbl paper.blg paper.pdf
 pdflatex paper && bibtex paper && pdflatex paper && pdflatex paper

### 2022-06-04 paper/paper.tex
Add total row to the table.
Order the citations.
Add ACM demo.

@@
-\documentclass[11pt,conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-\usepackage[sort,nocompress]{cite}
+\documentclass[11pt,sigplan,screen,nonacm]{acmart}
+% \documentclass[11pt,conference]{IEEEtran}
+% \IEEEoverridecommandlockouts
+% \usepackage[sort,nocompress]{cite}
+\usepackage{natbib}
+\setcitestyle{square, comma, numbers,sort&compress, super}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
 
@@
 
-\author{\IEEEauthorblockN{Benjamin Marks}
-  \IEEEauthorblockA{\textit{Stanford University} \\
-  benmarks@stanford.edu}
-  \and
-  \IEEEauthorblockN{Heron Yang}
-  \IEEEauthorblockA{\textit{Stanford University} \\
-  heronyang@stanford.edu}
-  \and
-  \IEEEauthorblockN{Yuewei Na}
-  \IEEEauthorblockA{\textit{Stanford University} \\
-  yueweina@stanford.edu}
+\author{Benjamin Marks}
+\affiliation{%
+  \institution{Stanford University}
+  \city{Stanford}
+  \state{California}
+  \country{USA}
 }
+\email{benmarks@stanford.edu}
 
-\maketitle
+\author{Heron Yang}
+\affiliation{%
+  \institution{Stanford University}
+  \city{Stanford}
+  \state{California}
+  \country{USA}
+}
+\email{heronyang@stanford.edu}
+
+\author{Yuewei Na}
+\affiliation{%
+  \institution{Stanford University}
+  \city{Stanford}
+  \state{California}
+  \country{USA}
+}
+\email{yueweina@stanford.edu}
+
+% \author{\IEEEauthorblockN{Benjamin Marks}
+  % \IEEEauthorblockA{\textit{Stanford University} \\
+  % benmarks@stanford.edu}
+  % \and
+  % \IEEEauthorblockN{Heron Yang}
+  % \IEEEauthorblockA{\textit{Stanford University} \\
+  % heronyang@stanford.edu}
+  % \and
+  % \IEEEauthorblockN{Yuewei Na}
+  % \IEEEauthorblockA{\textit{Stanford University} \\
+  % yueweina@stanford.edu}
+% }
 
@@
 
-\begin{IEEEkeywords}
-  Atomic Commit, Blockchain, Smart Contract, Distributed System, Database Transaction.
-\end{IEEEkeywords}
+\maketitle
+
+% \begin{IEEEkeywords}
+  % Atomic Commit, Blockchain, Smart Contract, Distributed System, Database Transaction.
+% \end{IEEEkeywords}
+\keywords{atomic commit, blockchain, smart contract, distributed system, database transaction}
 
@@
 
-We define a transaction as a list of operations starting with \texttt{Begin} followed by multiple \texttt{Get} or \texttt{Put} and ending with \texttt{Commit}.
+We define a transaction as a list of operations starting with \texttt{Begin()} followed by multiple \texttt{Get()} or \texttt{Put()} and ending with \texttt{Commit()}.
 The client creates a transaction and requests the coordinator to commit.
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{unsrt}
 \bibliography{references}
@@
     \hline
+    Total &169405 &0.000169405 &0.330\\
+    \hline
   \end{tabular}

### 2022-06-04 paper/paper.tex
Update..

@@
 \usepackage{amsmath,amssymb,amsfonts}
+\usepackage[toc,page]{appendix}
 \usepackage{algorithmic}
+\usepackage{fancyhdr}
 \usepackage{graphicx}
@@
   \item \texttt{StartVoting(transaction\_id, cohorts, vote\_timeout\_time)}\\Called by the coordinator to start voting on a transaction.
-\texttt{cohorts} is the number of cohorts that will vote on this transaction later.
-\texttt{vote\_timeout\_time} is the absolute time on the blockchain when the 2PC stops taking votes.
+\texttt{cohorts} is the number of cohorts that will vote on this transaction later.\\\texttt{vote\_timeout\_time} is the absolute time on the blockchain when the 2PC stops taking votes.
   \item \texttt{Vote(transaction\_id, cohort\_id, ballot)}\\Called by the cohort to vote \texttt{COMMIT} or \texttt{ABORT} on a transaction before timeout.
@@
 \begin{itemize}
-  \item \texttt{PrepareTransaction(\\PrepareTransactionRequest, PrepareTransactionResponse)}
-  \item \texttt{GetTransactionResult(\\GetTransactionResultRequest, GetTransactionResultResponse)}
+  \item \texttt{PrepareTransaction(\\PrepareTransactionRequest,\\PrepareTransactionResponse)}
+  \item \texttt{GetTransactionResult(\\GetTransactionResultRequest,\\GetTransactionResultResponse)}
 \end{itemize}
@@
 
-While testing in the Truffle development, we learned the gas used for successfully committing a transaction with two cohorts is around 0.33 USD \cite{gwei} (see \ref{appendix} for breakdowns).
+While testing in the Truffle development, we learned the gas used for successfully committing a transaction with two cohorts is around 0.33 USD \cite{gwei} (see Appendix \ref{appendix:blockchaingas} for breakdowns).
 
@@
 
-\appendix
-\label{appendix}
+% \appendix
+% \begin{table}[hbtp]
+  % \centering
+  % \caption{Gas used, cost in ETH/USD per blockchain transaction for a successful 2PC commit with 2 cohorts. Data as of 2022/05/30.}
+  % \label{table:cost}
+  % \begin{tabular}{|p{1.8cm}|p{1.8cm}|p{2cm}|p{1.8cm}|}
+    % \hline
+          % & Gas Used (Gwei) &  Cost in ETH & Cost in USD \\
+          % \hline
+    % StartVoting &73441 &0.000073441 & 0.143\\
+    % \hline
+    % Vote 1 &56451	&0.000056451 &0.110\\
+    % \hline
+    % Vote 2 &39513 &0.000039513 &0.077\\
+    % \hline
+    % Total &169405 &0.000169405 &0.330\\
+    % \hline
+  % \end{tabular}
+% \label{blockchaingas}
+% \end{table}
+
+\begin{appendices}
+
+  \section{Blockchain Gas Used}
+  \label{appendix:blockchaingas}
+
   \begin{table}[hbtp]
@@
     \label{table:cost}
-  \begin{tabular}{|l|c|c|c|}
+    \begin{tabular}{|p{1.8cm}|p{1.8cm}|p{2cm}|p{1.8cm}|}
       \hline
@@
   \end{table}
+
+\end{appendices}
+
 \end{document}

### 2022-06-04 paper/paper.tex
Fixed two issues.

@@
 The source code of our implementation is available at licensed under GNU Lesser General Public License.
-Our implementation depends on the following libraries: truffle \cite{github_truffle}, gganacheanache \cite{github_ganache}, sqlite3 \cite{github_sqlite3}, grpc \cite{github_grpc}, web3 \cite{github_web3}, abseil-cpp \cite{github_absl}, rule\_nodejs \cite{github_rulejs}, gflags \cite{github_gflags}, glog \cite{github_glog}, openssl \cite{github_openssl}, threadpool \cite{Shoshany2021_ThreadPool}, lmdb \cite{github_lmdb}, and lmdbxx \cite{github_lmdbxx}.
+Our implementation depends on the following libraries: truffle \cite{github_truffle}, ganache \cite{github_ganache}, sqlite3 \cite{github_sqlite3}, grpc \cite{github_grpc}, web3 \cite{github_web3}, abseil-cpp \cite{github_absl}, rule\_nodejs \cite{github_rulejs}, gflags \cite{github_gflags}, glog \cite{github_glog}, openssl \cite{github_openssl}, threadpool \cite{Shoshany2021_ThreadPool}, lmdb \cite{github_lmdb}, and lmdbxx \cite{github_lmdbxx}.
 

### 2022-06-04 paper/paper.tex
Update citation, lgpl, etc.

@@
 Then, we illustrate the implementation of individual components.
-The source code of our implementation is available at licensed under GNU Lesser General Public License.
+The source code of our implementation is available at licensed under GNU Lesser General Public License \cite{lgpl}.
 Our implementation depends on the following libraries: truffle \cite{github_truffle}, ganache \cite{github_ganache}, sqlite3 \cite{github_sqlite3}, grpc \cite{github_grpc}, web3 \cite{github_web3}, abseil-cpp \cite{github_absl}, rule\_nodejs \cite{github_rulejs}, gflags \cite{github_gflags}, glog \cite{github_glog}, openssl \cite{github_openssl}, threadpool \cite{Shoshany2021_ThreadPool}, lmdb \cite{github_lmdb}, and lmdbxx \cite{github_lmdbxx}.
@@
 Once we deployed our smart contract, the coordinator and the cohorts interacted with the contract through a smart contract client library.
-However, we didn’t find a proper C++ smart contract client library to use, so we used a JavaScript library (web.js[ref]) instead and added an additional gRPC layer to translate to C++.
+However, we didn’t find a proper C++ smart contract client library to use, so we used a JavaScript library (web.js \cite{github_web3}) instead and added an additional gRPC layer to translate to C++.
 

### 2022-06-04 paper/paper.tex
Fix typo.

@@
 Once we deployed our smart contract, the coordinator and the cohorts interacted with the contract through a smart contract client library.
-However, we didn’t find a proper C++ smart contract client library to use, so we used a JavaScript library (web.js \cite{github_web3}) instead and added an additional gRPC layer to translate to C++.
+However, we didn’t find a proper C++ smart contract client library to use, so we used a JavaScript library (web3.js \cite{github_web3}) instead and added an additional gRPC layer to translate to C++.
 

### 2022-06-04 paper/paper.tex
Fix few issues.

@@
 \begin{itemize}
-  \item \texttt{CommitAtomicTransaction( CommitAtomicTransactionRequest, CommitAtomicTransactionResponse)}
-  \item \texttt{GetTransactionResult( GetTransactionResultRequest, GetTransactionResultResponse)}
+  \item \texttt{CommitAtomicTransaction(\\CommitAtomicTransactionRequest,\\CommitAtomicTransactionResponse)}
+  \item \texttt{GetTransactionResult(\\GetTransactionResultRequest,\\GetTransactionResultResponse)}
 \end{itemize}
@@
       \item If the blockchain states that the transaction should be committed, it will call \texttt{DatabaseAdapter::Commit()} and release any acquired locks.
-  \item If the blockchain states that the transaction should be aborted, it will call \texttt{DatabaseAdapter::Abort()} and release any acquired locks.
+  \item If the blockchain states that the transaction should be aborted, it will call \\ \texttt{DatabaseAdapter::Abort()} and release any acquired locks.
   \end{enumerate}

### 2022-06-04 paper/paper.tex
Fix overfloating.

@@
 \begin{itemize}
-  \item \texttt{StartVoting(transaction\_id, cohorts, vote\_timeout\_time)}\\Called by the coordinator to start voting on a transaction.
+  \item \texttt{StartVoting(transaction\_id, cohorts,\\ vote\_timeout\_time)}\\Called by the coordinator to start voting on a transaction.
 \texttt{cohorts} is the number of cohorts that will vote on this transaction later.\\\texttt{vote\_timeout\_time} is the absolute time on the blockchain when the 2PC stops taking votes.
-  \item \texttt{Vote(transaction\_id, cohort\_id, ballot)}\\Called by the cohort to vote \texttt{COMMIT} or \texttt{ABORT} on a transaction before timeout.
+  \item \texttt{Vote(transaction\_id, cohort\_id,\\ballot)}\\Called by the cohort to vote \texttt{COMMIT} or \texttt{ABORT} on a transaction before timeout.
   \item \texttt{GetVotingDecision(transaction\_id)}\\Called by any component to check the voting state of a transaction.
@@
   \item Start a transaction with \texttt{DatabaseAdapter:: Begin()}.
-  \item Acquire the relevant locks for the sub-transaction.
+  \item Acquire the relevant locks for the sub-transa-ction.
   \item Execute a series of requested \texttt{Put} or \texttt{Get} operations in the sub-transaction. 
@@
 The coordinator returns a transaction ID for the client to look up the transaction result later.
-  \item The coordinator asks the blockchain to trace votes for the new transaction through \texttt{StartVoting()}.
+  \item The coordinator asks the blockchain to trace votes for the new transaction through \texttt{Start-Voting()}.
   \item The coordinator decomposes the transaction into sub-transactions and distributes them to the corresponding cohort through \texttt{PrepareTransaction()}.
@@
 \subsubsection{Blockchain}
-We wrote our two-phase-commit smart contract with Solidity and developed it with Truffle \cite{github_truffle}.
+We wrote our two-phase-\\commit smart contract with Solidity and developed it with Truffle \cite{github_truffle}.
 Truffle provided the development environment for us to compile smart contracts, test it, and deploy it locally for debugging.
@@
 \subsubsection{Database}
-We designed the \texttt{DatabaseAdapter} interface and implemented the methods with LMDB primitives.
+We designed the \texttt{DatabaseAda-pter} interface and implemented the methods with LMDB primitives.
 LMDB provides process-oriented functions like \texttt{open()}, \texttt{begin()}, \texttt{commit()}, \texttt{put()}, \texttt{get()}.

### 2022-06-04 paper/paper.tex
Fine tune.
Final edits.

@@
 \documentclass[11pt,sigplan,screen,nonacm]{acmart}
-% \documentclass[11pt,conference]{IEEEtran}
-% \IEEEoverridecommandlockouts
-% \usepackage[sort,nocompress]{cite}
 \usepackage{natbib}
@@
 
-% \author{\IEEEauthorblockN{Benjamin Marks}
-  % \IEEEauthorblockA{\textit{Stanford University} \\
-  % benmarks@stanford.edu}
-  % \and
-  % \IEEEauthorblockN{Heron Yang}
-  % \IEEEauthorblockA{\textit{Stanford University} \\
-  % heronyang@stanford.edu}
-  % \and
-  % \IEEEauthorblockN{Yuewei Na}
-  % \IEEEauthorblockA{\textit{Stanford University} \\
-  % yueweina@stanford.edu}
-% }
-
 \begin{abstract}
@@
 
+\settopmatter{printfolios=true} % add page number
+
+% hacky way to get some space below the author row.
+\begin{teaserfigure}
+\end{teaserfigure}
+
 \maketitle
 
-% \begin{IEEEkeywords}
-  % Atomic Commit, Blockchain, Smart Contract, Distributed System, Database Transaction.
-% \end{IEEEkeywords}
 \keywords{atomic commit, blockchain, smart contract, distributed system, database transaction}
@@
 The stateless coordinator can easily recover as long as it knows the participating cohorts.
-Second, The coordinator is stateless and horizontally scalable.
+Second, the coordinator is stateless and horizontally scalable.
 We can keep multiple jobs running concurrently and place a load-balancing layer on top of it to be tolerant of individual coordinator node failures.
@@
 \section{Design Overview}
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{./diagram.pdf}
+  \caption{This diagram displays the main components within our architecture and example steps of successfully committing a transaction.}
+  \label{fig:flow}
+\end{figure*}
+
 \subsection{Architecture and Components} \label{components}
@@
 We describe the details of each component in the remaining \ref{components}, the complete flow of handling a successful transaction in \ref{txnflow}, and how we handle failures in \ref{failurecases}.
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=\textwidth]{./diagram.pdf}
-  \caption{This diagram displays the main components within our architecture and example steps of successfully committing a transaction.}
-  \label{fig:Panel}
-\end{figure*}
 
@@
 \subsubsection{Database} \label{database}
-The database component is responsible for committing sub-transactions in each cohort and abstracts typical ACID databases with 5 APIs in a \texttt{DatabaseAdapter:} 
+The database component is responsible for committing sub-transactions in each cohort and abstracts typical ACID databases with five APIs in a \texttt{DatabaseAdapter:} 
 \begin{itemize}
@@
   \item \texttt{Abort()} aborts a started transaction.
-  \item \texttt{Put(key, value)} puts value in a row whose primary key is key.
-  \item \texttt{Get(key)} gets the value from the row whose primary key is key.
+  \item \texttt{Put(key, value)} puts value in a row whose primary key is \texttt{key}.
+  \item \texttt{Get(key)} gets the value from the row whose primary key is \texttt{key}.
 \end{itemize}
@@
 
-This section describes the end-to-end life of a transaction without handling failures.
+This section and Figure \ref{fig:flow} describe the end-to-end life of a transaction without failures.
 
@@
   \item The coordinator decomposes the transaction into sub-transactions and distributes them to the corresponding cohort through \texttt{PrepareTransaction()}.
-  \item Each cohort executes a sub-transaction by acquiring locks and forwarding the operations to the database up to the point before Commit().
+  \item Each cohort executes a sub-transaction by acquiring locks and forwarding the operations to the database up to the point before \texttt{Commit()}.
 The cohort writes an entry to the redo log to recover in case it crashes.
@@
 Then, we illustrate the implementation of individual components.
-The source code of our implementation is available at licensed under GNU Lesser General Public License \cite{lgpl}.
+The source code of our implementation is available at \cite{our_code} licensed under GNU Lesser General Public License \cite{lgpl}.
 Our implementation depends on the following libraries: truffle \cite{github_truffle}, ganache \cite{github_ganache}, sqlite3 \cite{github_sqlite3}, grpc \cite{github_grpc}, web3 \cite{github_web3}, abseil-cpp \cite{github_absl}, rule\_nodejs \cite{github_rulejs}, gflags \cite{github_gflags}, glog \cite{github_glog}, openssl \cite{github_openssl}, threadpool \cite{Shoshany2021_ThreadPool}, lmdb \cite{github_lmdb}, and lmdbxx \cite{github_lmdbxx}.
@@
 
-% \appendix
-% \begin{table}[hbtp]
-  % \centering
-  % \caption{Gas used, cost in ETH/USD per blockchain transaction for a successful 2PC commit with 2 cohorts. Data as of 2022/05/30.}
-  % \label{table:cost}
-  % \begin{tabular}{|p{1.8cm}|p{1.8cm}|p{2cm}|p{1.8cm}|}
-    % \hline
-          % & Gas Used (Gwei) &  Cost in ETH & Cost in USD \\
-          % \hline
-    % StartVoting &73441 &0.000073441 & 0.143\\
-    % \hline
-    % Vote 1 &56451	&0.000056451 &0.110\\
-    % \hline
-    % Vote 2 &39513 &0.000039513 &0.077\\
-    % \hline
-    % Total &169405 &0.000169405 &0.330\\
-    % \hline
-  % \end{tabular}
-% \label{blockchaingas}
-% \end{table}
-
 \begin{appendices}
@@
     \label{table:cost}
-    \begin{tabular}{|p{1.8cm}|p{1.8cm}|p{2cm}|p{1.8cm}|}
+    \begin{tabular}{|p{1.8cm}|p{1.8cm}|p{2cm}|p{1.5cm}|}
       \hline

### 2022-06-06 demo/start_blockchain.sh
Add demo for presentation.

@@ -1,0 +1,1 @@
+ bazel run //src/blockchain:deploy_ganache_command

### 2022-06-06 demo/start_coordinator.sh

@@ -1,0 +1,1 @@
+bazel run //src/coordinator:start_coordinator_with_blockchain_adapter_server -- --adapter_server_port=58600 -- --port=58000

### 2022-06-06 src/blockchain/client/adapter_server.js

@@ -57,7 +57,6 @@ function vote(call, callback) {
 }
 
 function getVotingDecision(call, callback) {
-  console.log('Received: getVotingDecision', call.request);
   contractClient.getVotingDecision(call.request.transaction_id, (result) => {
     console.log('Decision', result);
     var decision = 'VOTING_DECISION_UNKNOWN';
@@ -93,7 +92,6 @@ function main() {
     getVotingDecision: getVotingDecision,
     getHeartBeat: getHeartBeat
   });
-  console.log('Starting server')
   server.bindAsync(
       '0.0.0.0:' + port, grpc.ServerCredentials.createInsecure(), () => {
         server.start();

### 2022-06-06 src/blockchain/client/run_adapter_server.sh

@@ -25,7 +25,7 @@ $compile_script $1 $2
 shift
 shift
 adapter_server=$(rlocation __main__/src/blockchain/client/adapter_server_command.bash)
-contract_address=''
+contract_address='0xCfEB869F69431e42cdB54A4F4f105C19C080A601'
 adapter_server_port=61362
 # Don't change this since it's the first one in the deterministic ganache client.
 account='0x90F8bf6A479f320ead074411a4B0e7944Ea8c9C1'

### 2022-06-06 src/blockchain/deploy_ganache_without_console.sh

@@ -1,6 +1,5 @@
 #!/bin/bash
 set -eEuo pipefail
-
 # --- begin runfiles.bash initialization ---
 if [[ ! -d "${RUNFILES_DIR:-/dev/null}" && ! -f "${RUNFILES_MANIFEST_FILE:-/dev/null}" ]]; then
     if [[ -f "$0.runfiles_manifest" ]]; then
@@ -22,13 +21,5 @@ else
 fi
 # --- end runfiles.bash initialization ---
 
-setup_truffle="$(rlocation __main__/src/blockchain/setup_truffle.sh)"
-source $setup_truffle
-ganache="$(rlocation __main__/$3)"
-port=${4:-7545}
-# Kill the existing server if there is one.
-# Note that it takes about 1 second for it to be killed.
-fuser -k $port/tcp && sleep 1s || :
-$ganache --server.port=$port --wallet.deterministic --chain.networkId=5777 2>&1 | tee /tmp/ganache_log.txt &
-$truffle migrate
-sleep 5m
+deploy_ganache="$(rlocation __main__/src/blockchain/deploy_ganache.sh)"
+source $deploy_ganache

### 2022-06-06 src/client/client_main.cc

@@ -6,6 +6,7 @@
 #include "absl/container/flat_hash_map.h"
 #include "absl/container/flat_hash_set.h"
 #include "absl/flags/flag.h"
+#include "absl/flags/parse.h"
 #include "absl/status/statusor.h"
 #include "absl/strings/escaping.h"
 #include "absl/strings/str_cat.h"
@@ -22,7 +23,7 @@
 
 ABSL_FLAG(std::string, coordinator_port, "58000",
           "Port to send requests to the coordinator");
-ABSL_FLAG(std::string, request_file_prefix, "two_namespaces_simple",
+ABSL_FLAG(std::string, request_file_prefix, "",
           "Prefix of filename with requests to send to the coordinator");
 
 struct RunInfo {
@@ -206,8 +207,13 @@ void SendRequests(
                           grpc::InsecureChannelCredentials()));
   client.WaitForCoordinator();
   std::vector<std::future<client::ResponseOrStatus>> futures;
+  bool first_request = true;
   for (std::pair<coordinator::CommitAtomicTransactionRequest, bool>& request :
        requests) {
+    if (!first_request) {
+      std::this_thread::sleep_for(absl::ToChronoSeconds(absl::Seconds(5)));
+    }
+    first_request = false;
     if (request.first.config().has_presumed_abort_time()) {
       request.first.mutable_config()
           ->mutable_presumed_abort_time()
@@ -225,8 +231,9 @@ void SendRequests(
   }
 }
 
-int main(int /*argc*/, char** /*argv*/) {
+int main(int argc, char** argv) {
   try {
+    absl::ParseCommandLine(argc, argv);
     absl::StatusOr<RunInfo> run_info = CreateRunInfoFromFile(
         absl::StrCat("src/client/requests/",
                      absl::GetFlag(FLAGS_request_file_prefix), ".textproto"));
